### 2024-07-15 ProxySelection.Rmd
results

---
title: "Step1_LoadData"
author: "Yang Lei"
date: "2024-07-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(autoCovariateSelection)
require(dplyr)
require(cobalt)
require(WeightIt)
library(tidyverse)
library(ggplot2)
library(lmtest)
library(MASS)
library(Boruta)
library(GA)
library(mclust)
library(penalizedSVM)
library(xgboost)
library(caret)
require(autoCovariateSelection)
```

**Codechuck 0. Loading data -- Generating Recurrence Covariates**

_Step 0: Analytic data, Step 1: Proxy sources, Step 2: Emperical, Step 3: Recurrence_
```{r loading-data, warning=FALSE, cache=TRUE}
# ------------------- Select proxy based on the results from out2
# ------------------- change the code after Step 3: Recurrence
load("data/analytic3cycles.RData")

analytic <- data.complete
idx <- analytic$id
outcome <- as.numeric(analytic$diabetes == "Yes") 
exposure <- as.numeric(analytic$obese == "Yes")

domain <- "dx"
analytic.dfx <- as.data.frame(cbind(idx, exposure, outcome, domain))
dat.proxy.long <- subset(dat.proxy.long, 
                         icd10 != "E66") # Overweight and obesity
dat.proxy.long <- subset(dat.proxy.long, 
                         icd10 != "O24") # Gestational diabetes mellitus
dat.proxy.long <- subset(dat.proxy.long, 
                         icd10 != "E10") # Type 1 diabetes mellitus
dat.proxy.long <- subset(dat.proxy.long, 
                         icd10 != "E11") # Type 2 diabetes mellitus
proxy.var.long <- dat.proxy.long
proxy.var.long$idx <- proxy.var.long$id
proxy.var.long$id <- NULL
dfx <- merge(analytic.dfx, proxy.var.long, by = "idx")
# ------------------- now the proxy data was combined to the A and Y

# basetable <- dfx %>% select(idx, exposure, outcome) %>% distinct()
basetable <- dfx[, c(1:3)] %>% distinct()
patientIds <- basetable$idx
length(patientIds)
# ------------------- number of observations

save(dfx, patientIds, basetable, file = "data/dfx.RData")
load(file = "data/dfx.RData")

step1 <- get_candidate_covariates(df = dfx,  
                                  domainVarname = "domain",
                                  eventCodeVarname = "icd10", 
                                  patientIdVarname = "idx",
                                  patientIdVector = patientIds,
                                  n = 200, 
                                  min_num_patients = 20)
out1 <- step1$covars_data
# ------------------- top 200 prevalent icd10 with min 20 patients

step2 <- get_recurrence_covariates(df = out1, 
                                   patientIdVarname = "idx",
                                   eventCodeVarname = "icd10", 
                                   patientIdVector = patientIds)
out2 <- step2$recurrence_data

exposure <- "obese"
outcome <- "diabetes" 
investigator.specified.covariates <- 
  c(# Demographic
    "age.cat", "sex", "education", "race", 
    "marital", "income", "born", "year",

    # health history related variables/access
    "diabetes.family.history", "medical.access",

    # behavioral
    "smoking", "diet.healthy", "physical.activity", "sleep",

    # Laboratory 
    "uric.acid", "protein.total", "bilirubin.total", "phosphorus",
    "sodium", "potassium", "globulin", "calcium.total", 
    "systolicBP", "diastolicBP", "high.cholesterol"
  )
```

**Method 1: Bross Formula**
```{r, warning=FALSE, cache=TRUE}
out3 <- get_prioritised_covariates(df = out2,
                                   patientIdVarname = "idx", 
                                   exposureVector = basetable$exposure,
                                   outcomeVector = basetable$outcome,
                                   patientIdVector = patientIds, 
                                   k = 100)
hdps.dim <- out3$autoselected_covariate_df

load("data/analytic3cycles.RData")
hdps.dim$id <- hdps.dim$idx
hdps.dim$idx <- NULL
hdps.data <- merge(data.complete[,c("id",
                                    outcome, 
                                    exposure, 
                                    investigator.specified.covariates)], 
                   hdps.dim, by = "id")

hdps.data$exposure <- as.numeric(I(hdps.data$obese=='Yes'))
hdps.data$outcome <- as.numeric(I(hdps.data$diabetes=='Yes'))

proxy.list.sel <- names(out3$autoselected_covariate_df[,-1])
proxyform <- paste0(proxy.list.sel, collapse = "+")
covform <- paste0(investigator.specified.covariates, collapse = "+")
rhsformula <- paste0(c(covform, proxyform), collapse = "+")
ps.formula <- as.formula(paste0("exposure", "~", rhsformula))

W.out <- weightit(ps.formula, 
                  data = hdps.data, 
                  estimand = "ATE",
                  method = "ps")
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
fit <- glm(out.formula,
           data = hdps.data,
           weights = W.out$weights,
           family= binomial(link = "logit"))
fit.summary <- summary(fit)$coef["exposure",
                                 c("Estimate", 
                                   "Std. Error", 
                                   "Pr(>|z|)")]
fit.ci <- confint(fit, level = 0.95)["exposure", ]
fit.summary_with_ci <- c(fit.summary, fit.ci)
fit.summary_with_ci

fs.bross <- proxy.list.sel
sum_ci.bross <- c(exp(fit.summary_with_ci[1]), fit.summary_with_ci)
names(sum_ci.bross) <- c("Odds Ratio",
                          "Estimate",
                          "Std. Error",
                          "Pr(>|z|)",
                          "2.5 %",
                          "97.5 %")

print("No need to print the number of proxy variables selected, as it is set to be 100.")

print("The Odds Ratio by Bross Formula")
round(sum_ci.bross, 3)
```

**Method 2: Hybrid Bross Formula & Lasso**
```{r, warning=FALSE, cache=TRUE}
set.seed(42)
proxy.list <- names(out3$autoselected_covariate_df[,-1])
covarsTfull <- c(investigator.specified.covariates, proxy.list)
Y.form <- as.formula(paste0(c("outcome~ exposure", 
                              covarsTfull), collapse = "+") )
covar.mat <- model.matrix(Y.form, data = hdps.data)[,-1]
lasso.fit<-glmnet::cv.glmnet(y = hdps.data$outcome, 
                             x = covar.mat, 
                             type.measure='mse',
                             family="binomial",
                             alpha = 1, 
                             nfolds = 5)
coef.fit<-coef(lasso.fit,s='lambda.min',exact=TRUE)
sel.variables<-row.names(coef.fit)[which(as.numeric(coef.fit)!=0)]
proxy.list.sel.hybrid <- proxy.list[proxy.list %in% sel.variables]
length(proxy.list.sel.hybrid)
fs.hybrid <- proxy.list.sel.hybrid

proxyform <- paste0(proxy.list.sel.hybrid, collapse = "+")
rhsformula <- paste0(c(covform, proxyform), collapse = "+")
ps.formula <- as.formula(paste0("exposure", "~", rhsformula))

W.out <- weightit(ps.formula, 
                  data = hdps.data, 
                  estimand = "ATE",
                  method = "ps")

out.formula <- as.formula(paste0("outcome", "~", "exposure"))
fit <- glm(out.formula,
           data = hdps.data,
           weights = W.out$weights,
           family= binomial(link = "logit"))
fit.summary <- summary(fit)$coef["exposure",
                                 c("Estimate", 
                                   "Std. Error", 
                                   "Pr(>|z|)")]
fit.ci <- confint(fit, level = 0.95)["exposure", ]
fit.summary_with_ci.h <- c(fit.summary, fit.ci)
sum_ci.hybrid <- c(exp(fit.summary_with_ci.h[1]), fit.summary_with_ci.h)

paste("The number of proxy variables selected:", length(proxy.list.sel.hybrid))
print("All proxy variables selected are in common as they are all selected from 
      the proxy variable list by the bross formula")

print("The Odds Ratio by Bross Formula & Lasso")
round(sum_ci.hybrid, 3)
```

_Before implementing other methods, drop the information from method 1 (Bross formula). Continue with the results from Step 3: Recurrence (table `out2`)._
```{r, warning=FALSE, cache=TRUE}
hdps.dim <- out2

load("data/analytic3cycles.RData")
hdps.dim$id <- hdps.dim$idx
hdps.dim$idx <- NULL
hdps.data <- merge(data.complete[,c("id",
                                    outcome, 
                                    exposure, 
                                    investigator.specified.covariates)], 
                   hdps.dim, by = "id")

hdps.data$exposure <- as.numeric(I(hdps.data$obese=='Yes'))
hdps.data$outcome <- as.numeric(I(hdps.data$diabetes=='Yes'))

proxy.list.sel <- names(out2[,-1])
proxyform <- paste0(proxy.list.sel, collapse = "+")
covform <- paste0(investigator.specified.covariates, collapse = "+")
rhsformula <- paste0(c(covform, proxyform), collapse = "+")
ps.formula <- as.formula(paste0("exposure", "~", rhsformula))

proxy.list <- names(out2[,-1])    # length(proxy.list) :  142
covarsTfull <- c(investigator.specified.covariates, proxy.list)
Y.form <- as.formula(paste0(c("outcome~ exposure", 
                              covarsTfull), collapse = "+") )

```

**Method 3: Pure Lasso**
```{r, warning=FALSE, cache=TRUE}
set.seed(42) # different seed -> different length for the selected proxy list
covar.mat <- model.matrix(Y.form, data = hdps.data)[,-1]
lasso.fit <- glmnet::cv.glmnet(y = hdps.data$outcome, 
                               x = covar.mat, 
                               type.measure='mse',
                               family="binomial",
                               alpha = 1, 
                               nfolds = 5)
coef.fit<-coef(lasso.fit,s='lambda.min',exact=TRUE)
sel.variables<-row.names(coef.fit)[which(as.numeric(coef.fit)!=0)]
proxy.list.sel.ml <- proxy.list[proxy.list %in% sel.variables]
length(proxy.list.sel.ml)
fs.lasso <- proxy.list.sel.ml
proxy.lasso <- paste0(fs.lasso, collapse = "+")
form.lasso <- paste0(c(covform, proxy.lasso), collapse = "+")
formula.lasso <- as.formula(paste0("exposure", "~", form.lasso))

W.out.lasso <- weightit(formula.lasso, 
                        data = hdps.data, 
                        estimand = "ATE",
                        method = "ps")

hdps.data$ps.l <- W.out.lasso$ps
hdps.data$w.l <- W.out.lasso$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.lasso <- glm(out.formula,
                 data = hdps.data,
                 weights = W.out.lasso$weights,
                 family= binomial(link = "logit"))
summary.lasso <- summary(fit.lasso)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.lasso <- confint(fit.lasso, level = 0.95)["exposure", ]
sum_ci.lasso <- c(exp(summary.lasso[1]), summary.lasso, ci.lasso)
names(sum_ci.lasso) <- c("Odds Ratio",
                         "Estimate",
                         "Std. Error",
                         "Pr(>|z|)",
                         "2.5 %",
                         "97.5 %")

# Q1: How many proxy variables are in common?
# length(fs.lasso)  # 60 features selected by lasso
# length(fs.lasso[fs.lasso %in% fs.bross])  # 47 of 60 also selected by bross
print("How many proxy variables are in common?")
paste("There are", length(fs.lasso), "features selected by lasso")
paste(length(fs.lasso[fs.lasso %in% fs.bross]), "of the", length(fs.lasso), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.lasso[!(fs.lasso %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Method Pure Lasso")
sum_ci.lasso
```

**Method 4: Elastic Net**
```{r, warning=FALSE, cache=TRUE}
set.seed(42)
covar.mat <- model.matrix(Y.form, data = hdps.data)[,-1]
elas.fit <- glmnet::cv.glmnet(y = hdps.data$outcome, 
                             x = covar.mat, 
                             type.measure='mse',
                             family="binomial",
                             alpha = 0.5, 
                             nfolds = 5)
coef.fit<-coef(elas.fit, s='lambda.min', exact=TRUE)
sel.variables<-row.names(coef.fit)[which(as.numeric(coef.fit)!=0)]
proxy.list.sel.elastic <- proxy.list[proxy.list %in% sel.variables]
length(proxy.list.sel.elastic)
fs.elastic <- proxy.list.sel.elastic

proxy.elastic <- paste0(fs.elastic, collapse = "+")
form.elastic <- paste0(c(covform, proxy.elastic), collapse = "+")
formula.elastic <- as.formula(paste0("exposure", "~", form.elastic))

W.out.elastic <- weightit(formula.elastic, 
                       data = hdps.data, 
                       estimand = "ATE",
                       method = "ps")

hdps.data$ps.e <- W.out.elastic$ps
hdps.data$w.e <- W.out.elastic$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.elastic <- glm(out.formula,
                data = hdps.data,
                weights = W.out.elastic$weights,
                family= binomial(link = "logit"))
summary.elastic <- summary(fit.elastic)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.elastic <- confint(fit.elastic, level = 0.95)["exposure", ]
sum_ci.elastic <- c(exp(summary.elastic[1]), summary.elastic, ci.elastic)
names(sum_ci.elastic) <- c("Odds Ratio",
                         "Estimate",
                         "Std. Error",
                         "Pr(>|z|)",
                         "2.5 %",
                         "97.5 %")

# Q1: How many proxy variables are in common (Elastic Net vs Bross Formula)?
# length(fs.elastic)  # 69 features selected by elastic net
# length(fs.elastic[fs.elastic %in% fs.bross])  # 54 of 69 also selected by bross formula
print("How many proxy variables are in common?")
paste("There are", length(fs.elastic), "features selected by elas")
paste(length(fs.elastic[fs.elastic %in% fs.bross]), "of the", length(fs.elastic), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.elastic[!(fs.elastic %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Elastic Net")
sum_ci.elastic
```

**Method 5: XGBoost**
```{r, warning=FALSE, cache=TRUE}
xgb_fit <- xgboost(data = covar.mat, label = hdps.data$outcome, 
                    max.depth = 20, eta = 1, nthread = 2, nrounds = 5, 
                    objective = "binary:logistic")

xgb_imp <- xgb.importance(model = xgb_fit)
xgb.plot.importance(importance_matrix = xgb_imp)
fs.xgboost <- xgb_imp$Feature
fs.xgboost <- proxy.list[proxy.list %in% fs.xgboost]
proxy.xgboost <- paste0(fs.xgboost, collapse = "+")
form.xgboost <- paste0(c(covform, proxy.xgboost), collapse = "+")
formula.xgboost <- as.formula(paste0("exposure", "~", form.xgboost))

W.out.xgboost <- weightit(formula.xgboost, 
                          data = hdps.data, 
                          estimand = "ATE",
                          method = "ps")

hdps.data$ps.xgboost <- W.out.xgboost$ps
hdps.data$w.xgboost <- W.out.xgboost$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.xgboost <- glm(out.formula,
                   data = hdps.data,
                   weights = W.out.xgboost$weights,
                   family= binomial(link = "logit"))
summary.xgboost <- summary(fit.xgboost)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.xgboost <- confint(fit.xgboost, level = 0.95)["exposure", ]
sum_ci.xgboost <- c(exp(summary.xgboost[1]), summary.xgboost, ci.xgboost)
names(sum_ci.xgboost) <- c("Odds Ratio",
                           "Estimate",
                           "Std. Error",
                           "Pr(>|z|)",
                           "2.5 %",
                           "97.5 %")

# Q1: How many proxy variables are in common?
# length(fs.xgboost) 
# length(fs.xgboost[fs.xgboost %in% fs.bross])
print("How many proxy variables are in common?")
paste("There are", length(fs.xgboost), "features selected by xgboost")
paste(length(fs.xgboost[fs.xgboost %in% fs.bross]), "of the", length(fs.xgboost), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.xgboost[!(fs.xgboost %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by XGBoost")
sum_ci.xgboost
```

**Method 6: GA**

_GA (Genetic Algorithm) has no specific internal model building mechanism and would wrap around whatever model the user specifies_
```{r, warning=FALSE, cache=TRUE}
gaOpt <- function(vars, IV.train, DV.train) {
	varNames <- colnames(IV.train) #getting names of all variables
	selectedVarNames <- varNames[vars == "1"] # getting names of selected vars from GA
	gaSolutionData <- IV.train[,selectedVarNames] # keeping only those selected vars

	gaDat <- cbind(gaSolutionData,DV.train) # combining selected variables with outcome variable
	gaMod <- glm(DV.train ~ ., family = "binomial", data = gaDat) #build model
	gaProb <- predict(gaMod, IV.train, type = "response") # get probabilities
	gaPred <- ifelse(gaProb >= .8, 1, 0) # get predicted 0s and 1s

	ari <- adjustedRandIndex(gaPred, DV.train)
	return(ari)
}

ga.solution <- ga(fitness = function(vars)
 gaOpt(vars=vars, 
       IV.train=data.frame(covar.mat),
       DV.train=hdps.data$outcome),
 type = "binary", 
 nBits = ncol(covar.mat),
 names = colnames(covar.mat), 
 seed = 42,
 run=5)

allVarNames <- colnames(covar.mat)
sel.variables.GA <- allVarNames[ga.solution@solution[1,]==1]
fs.GA <- proxy.list[proxy.list %in% sel.variables.GA]
proxy.GA <- paste0(fs.GA, collapse = "+")
form.GA <- paste0(c(covform, proxy.GA), collapse = "+")
formula.GA <- as.formula(paste0("exposure", "~", form.GA))

W.out.GA <- weightit(formula.GA, 
                         data = hdps.data, 
                         estimand = "ATE",
                         method = "ps")

hdps.data$ps.ga <- W.out.GA$ps
hdps.data$w.ga <- W.out.GA$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.GA <- glm(out.formula,
                  data = hdps.data,
                  weights = W.out.GA$weights,
                  family= binomial(link = "logit"))
summary.GA <- summary(fit.GA)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.GA <- confint(fit.GA, level = 0.95)["exposure", ]
sum_ci.GA <- c(exp(summary.GA[1]), summary.GA, ci.GA)
names(sum_ci.GA) <- c("Odds Ratio",
                          "Estimate",
                          "Std. Error",
                          "Pr(>|z|)",
                          "2.5 %",
                          "97.5 %")

# Q1: How many proxy variables are in common (GA vs Bross Formula)?
# length(fs.GA)  # 64 features selected by GA
# length(fs.GA[fs.GA %in% fs.bross])  # 44 of 64 also selected by bross formula
print("How many proxy variables are in common?")
paste("There are", length(fs.GA), "features selected by GA")
paste(length(fs.GA[fs.GA %in% fs.bross]), "of the", length(fs.GA), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.GA[!(fs.GA %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by GA:")
sum_ci.GA
```

**Method 7: Elastic SCAD SVM**

_Elastic SCAD SVM uses an approach like LASSO and Elastic Net in that it selects variables via the addition of a penalty._

_(Long Running Time)_
```{r, warning=FALSE, cache=TRUE}
svmTrainOutcome <- hdps.data$outcome
svmTrainOutcome[svmTrainOutcome == 0] <- -1
svmTrainPreds <- data.frame(covar.mat)

bounds <- t(data.frame(log2lambda1=c(-10, 10), log2lambda2=c(-10,10)))
colnames(bounds)<-c("lower", "upper")
svm.model <- svmfs(x=covar.mat,
				   y = svmTrainOutcome,
				   fs.method = "scad+L2",
				   bounds=bounds,
				   grid.search = "interval",
				   inner.val.method = "cv",
				   show = "none",
				   parms.coding = "none",
				   seed=42)
```

```{r, warning=FALSE, cache=TRUE}
fs.svm <- colnames(covar.mat)[svm.model$model$fit.info$model.list$model$xind]
fs.svm <- proxy.list[proxy.list %in% fs.svm]
proxy.svm <- paste0(fs.svm, collapse = "+")
form.svm <- paste0(c(covform, proxy.svm), collapse = "+")
formula.svm <- as.formula(paste0("exposure", "~", form.svm))

W.out.svm <- weightit(formula.svm, 
                        data = hdps.data, 
                        estimand = "ATE",
                        method = "ps")

hdps.data$ps.svm <- W.out.svm$ps
hdps.data$w.svm <- W.out.svm$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.svm <- glm(out.formula,
                 data = hdps.data,
                 weights = W.out.svm$weights,
                 family= binomial(link = "logit"))
summary.svm <- summary(fit.svm)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.svm <- confint(fit.svm, level = 0.95)["exposure", ]
sum_ci.svm <- c(exp(summary.svm[1]), summary.svm, ci.svm)
names(sum_ci.svm) <- c("Odds Ratio",
                         "Estimate",
                         "Std. Error",
                         "Pr(>|z|)",
                         "2.5 %",
                         "97.5 %")

# Q1: How many proxy variables are in common?
# length(fs.svm) 
# length(fs.svm[fs.svm %in% fs.bross])
print("How many proxy variables are in common?")
paste("There are", length(fs.svm), "features selected by svm")
paste(length(fs.svm[fs.svm %in% fs.bross]), "of the", length(fs.svm), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.svm[!(fs.svm %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Elastic SCAD SVM")
sum_ci.svm
```

**Method 8: Boruta Selection**

_Boruta Selection would give the feature importance ranking based on the Random Forests algorithm which are able to capture some non-linear trends but not easy to interpret._

_(Long Running Time)_
```{r, warning=FALSE, cache=TRUE}
set.seed(42)
boruta_output <- Boruta(Y.form, data = hdps.data, doTrace = 1)
sel.variables.bor <- names(boruta_output$finalDecision)[which(boruta_output$finalDecision != "Rejected")]
fs.boruta <- proxy.list[proxy.list %in% sel.variables.bor]
proxy.boruta <- paste0(fs.boruta, collapse = "+")
form.boruta <- paste0(c(covform, proxy.boruta), collapse = "+")
formula.boruta <- as.formula(paste0("exposure", "~", form.boruta))

W.out.boruta <- weightit(formula.boruta, 
                         data = hdps.data, 
                         estimand = "ATE",
                         method = "ps")

hdps.data$ps.b <- W.out.boruta$ps
hdps.data$w.b <- W.out.boruta$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.boruta <- glm(out.formula,
                  data = hdps.data,
                  weights = W.out.boruta$weights,
                  family= binomial(link = "logit"))
summary.boruta <- summary(fit.boruta)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.boruta <- confint(fit.boruta, level = 0.95)["exposure", ]
sum_ci.boruta <- c(exp(summary.boruta[1]), summary.boruta, ci.boruta)
names(sum_ci.boruta) <- c("Odds Ratio",
                          "Estimate",
                          "Std. Error",
                          "Pr(>|z|)",
                          "2.5 %",
                          "97.5 %")

# Q1: How many proxy variables are in common (Boruta vs Bross Formula)?
# length(fs.boruta)  # 28 features selected by boruta selection
# length(fs.boruta[fs.boruta %in% fs.bross])  # 24 of 28 also selected by bross formula
print("How many proxy variables are in common?")
paste("There are", length(fs.boruta), "features selected by boruta")
paste(length(fs.boruta[fs.boruta %in% fs.bross]), "of the", length(fs.boruta), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.boruta[!(fs.boruta %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Boruta Selection")
sum_ci.boruta
```

**Method 9.1: Stepwise Regression (Forward)**
```{r, warning=FALSE, cache=TRUE}
set.seed(42)

initial_formula <- as.formula(paste0("outcome~exposure+",
                                     covform,
                                     collapse = "+"))
full_formula <- as.formula(paste0("outcome~exposure+",
                                  rhsformula,
                                  collapse = "+"))

initial_model <- glm(initial_formula, data = hdps.data, family = binomial)
full_model <- glm(full_formula, data = hdps.data, family = binomial)
stepwise_forward <- stepAIC(initial_model, 
                            scope = list(lower = initial_model, 
                                         upper = full_model), 
                            direction = "forward")

fs.forward <- all.vars(formula(stepwise_forward))[-1]
fs.forward <- proxy.list[proxy.list %in% fs.forward]
proxy.forward <- paste0(fs.forward, collapse = "+")
form.forward <- paste0(c(covform, proxy.forward), collapse = "+")
formula.forward <- as.formula(paste0("exposure", "~", form.forward))

W.out.forward <- weightit(formula.forward, 
                          data = hdps.data, 
                          estimand = "ATE",
                          method = "ps")

hdps.data$ps.forward <- W.out.forward$ps
hdps.data$w.forward <- W.out.forward$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.forward <- glm(out.formula,
                   data = hdps.data,
                   weights = W.out.forward$weights,
                   family= binomial(link = "logit"))
summary.forward <- summary(fit.forward)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.forward <- confint(fit.forward, level = 0.95)["exposure", ]
sum_ci.forward <- c(exp(summary.forward[1]), summary.forward, ci.forward)
names(sum_ci.forward) <- c("Odds Ratio",
                           "Estimate",
                           "Std. Error",
                           "Pr(>|z|)",
                           "2.5 %",
                           "97.5 %")

# Q1: How many proxy variables are in common?
# length(fs.forward) 
# length(fs.forward[fs.forward %in% fs.bross])
print("How many proxy variables are in common?")
paste("There are", length(fs.forward), "features selected by forward")
paste(length(fs.forward[fs.forward %in% fs.bross]), "of the", length(fs.forward), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.forward[!(fs.forward %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Stepwise Forward Selection")
sum_ci.forward
```

**Method 9.2: Stepwise Regression (Backward)**
```{r, warning=FALSE, cache=TRUE}
stepwise_backward <- stepAIC(full_model, direction = "backward")
fs.backward <- proxy.list[proxy.list %in% fs.backward]
fs.backward <- fs.backward[fs.backward != "exposure"]
proxy.backward <- paste0(fs.backward, collapse = "+")
form.backward <- paste0(c(covform, proxy.backward), collapse = "+")
formula.backward <- as.formula(paste0("exposure", "~", form.backward))

W.out.backward <- weightit(formula.backward, 
                          data = hdps.data, 
                          estimand = "ATE",
                          method = "ps")

hdps.data$ps.backward <- W.out.backward$ps
hdps.data$w.backward <- W.out.backward$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.backward <- glm(out.formula,
                   data = hdps.data,
                   weights = W.out.backward$weights,
                   family= binomial(link = "logit"))
summary.backward <- summary(fit.backward)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.backward <- confint(fit.backward, level = 0.95)["exposure", ]
sum_ci.backward <- c(exp(summary.backward[1]), summary.backward, ci.backward)
names(sum_ci.backward) <- c("Odds Ratio",
                           "Estimate",
                           "Std. Error",
                           "Pr(>|z|)",
                           "2.5 %",
                           "97.5 %")

# Q1: How many proxy variables are in common?
# length(fs.backward) 
# length(fs.backward[fs.backward %in% fs.bross])
print("How many proxy variables are in common?")
paste("There are", length(fs.backward), "features selected by backward")
paste(length(fs.backward[fs.backward %in% fs.bross]), "of the", length(fs.backward), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.backward[!(fs.backward %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Stepwise Backward Selection")
sum_ci.backward
```

**Method 9.3: Stepwise Regression (Both)**
```{r, warning=FALSE, cache=TRUE}
stepwise_both <- stepAIC(initial_model, 
                            scope = list(lower = initial_model, 
                                         upper = full_model), 
                            direction = "both")
fs.both <- all.vars(formula(stepwise_both))[-1]
fs.both <- proxy.list[proxy.list %in% fs.both]
proxy.both <- paste0(fs.both, collapse = "+")
form.both <- paste0(c(covform, proxy.both), collapse = "+")
formula.both <- as.formula(paste0("exposure", "~", form.both))

W.out.both <- weightit(formula.both, 
                          data = hdps.data, 
                          estimand = "ATE",
                          method = "ps")

hdps.data$ps.both <- W.out.both$ps
hdps.data$w.both <- W.out.both$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.both <- glm(out.formula,
                   data = hdps.data,
                   weights = W.out.both$weights,
                   family= binomial(link = "logit"))
summary.both <- summary(fit.both)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.both <- confint(fit.both, level = 0.95)["exposure", ]
sum_ci.both <- c(exp(summary.both[1]), summary.both, ci.both)
names(sum_ci.both) <- c("Odds Ratio",
                           "Estimate",
                           "Std. Error",
                           "Pr(>|z|)",
                           "2.5 %",
                           "97.5 %")

# Q1: How many proxy variables are in common?
# length(fs.both) 
# length(fs.both[fs.both %in% fs.bross])
print("How many proxy variables are in common?")
paste("There are", length(fs.both), "features selected by both")
paste(length(fs.both[fs.both %in% fs.bross]), "of the", length(fs.both), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.both[!(fs.both %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Stepwise Both Selection")
sum_ci.both
```

Comparing the Results:
```{r, warning=FALSE, cache=TRUE}
sum_ci.all <- rbind(sum_ci.bross, sum_ci.hybrid, sum_ci.lasso, sum_ci.elastic, 
                    sum_ci.xgboost, sum_ci.GA, sum_ci.svm, sum_ci.boruta,
                    sum_ci.forward, sum_ci.backward, sum_ci.both)

OR.data <- data.frame(method = c("bross", "hybrid", "lasso", "elastic", 
                                 "xgboost", "GA", "SVM", "boruta", 
                                 "forward", "backward", "both"),
                      OR = exp(sum_ci.all[,2]),
                      lower = exp(sum_ci.all[,5]),
                      upper = exp(sum_ci.all[,6]))

OR.data$method <- factor(OR.data$method, 
                         levels = c("both", "backward", "forward",
                                    "boruta", "SVM", "GA", "xgboost", 
                                    "elastic", "lasso", "hybrid", "bross"))

OR.ci_plot <- ggplot(OR.data, aes(y = method, x = OR)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.2) + 
  labs(title = "Odds Ratios with 95% Confidence Intervals",
       x = "Odds Ratio",
       y = "Method")

fs.all <- data.frame(numFeature = c(length(fs.bross), 
                                    length(fs.hybrid),
                                    length(fs.lasso), 
                                    length(fs.elastic), 
                                    length(fs.xgboost),
                                    length(fs.GA),
                                    length(fs.svm),
                                    length(fs.boruta),
                                    length(fs.forward),
                                    length(fs.backward),
                                    length(fs.both)),
                     numInCommon = c(NA,
                                     length(fs.hybrid[fs.hybrid %in% fs.bross]),
                                     length(fs.lasso[fs.lasso %in% fs.bross]),
                                     length(fs.elastic[fs.elastic %in% fs.bross]),
                                     length(fs.xgboost[fs.xgboost %in% fs.bross]),
                                     length(fs.GA[fs.GA %in% fs.bross]),
                                     length(fs.svm[fs.svm %in% fs.bross]),
                                     length(fs.boruta[fs.boruta %in% fs.bross]),
                                     length(fs.forward[fs.forward %in% fs.bross]),
                                     length(fs.backward[fs.backward %in% fs.bross]),
                                     length(fs.both[fs.both %in% fs.bross])))
rownames(fs.all) <-  c("bross", "hybrid", "lasso", "elastic", 
                       "xgb", "GA", "SVM", "boruta",
                       "forward", "backward", "both")

fs.all$pctCommon <- fs.all$numInCommon / fs.all$numFeature

fs.commonOneToOne <- data.frame(ini = rep(NA, 11))

for (i in 1:11) {
  col_name <- OR.data$method[i]
  if (col_name == "SVM") {
    col_name <- "svm"
  }
  na.i <- 1
  na <- c()
  while (na.i < i) {
    na <- c(na, NA)
    na.i <- na.i + 1
  }
  num_common <- c(length(fs.bross[fs.bross %in% get(paste0("fs.",col_name))]),
                  length(fs.hybrid[fs.hybrid %in% get(paste0("fs.",col_name))]),
                  length(fs.lasso[fs.lasso %in% get(paste0("fs.",col_name))]),
                  length(fs.elastic[fs.elastic %in% get(paste0("fs.",col_name))]),
                  length(fs.xgboost[fs.xgboost %in% get(paste0("fs.",col_name))]),
                  length(fs.GA[fs.GA %in% get(paste0("fs.",col_name))]),
                  length(fs.svm[fs.svm %in% get(paste0("fs.",col_name))]),
                  length(fs.boruta[fs.boruta %in% get(paste0("fs.",col_name))]),
                  length(fs.forward[fs.forward %in% get(paste0("fs.",col_name))]),
                  length(fs.backward[fs.backward %in% get(paste0("fs.",col_name))]),
                  length(fs.both[fs.both %in% get(paste0("fs.",col_name))]))
  fs.commonOneToOne[i] <- c(na, num_common[i:11])
}

colnames(fs.commonOneToOne) <- rownames(fs.all)
rownames(fs.commonOneToOne) <- rownames(fs.all)
```

Print the comparing plot & table:
```{r}
print("Comparison of odds ratios and the confidence intervals by different proxy selection methods:")
OR.ci_plot
```

```{r}
print("Comparison of the number & percentage of proxy variables selected by different methods in common with that by the bross formula:")
fs.all
```

```{r}
print("One-to-one comparison of the number of proxy variables selected by different methods")
fs.commonOneToOne
```