### 2024-07-31 ProxySelection.Rmd
Add estimated RD & the SE.
Updated / added code chunks: Chunk 4 (bross RD), Chunk 6 (hybrid RD), Chunk 9 (lasso RD), Chunk 11 (enet ED), Chunk 13 (xgb RD), Chunk 15 (ga RD), Chunk 18 (svm RD), Chunk 20 (Boruta RD), Chunk 22 (fs RD), Chunk 24 (bs RD), Chunk 26 (sw RD), Chunk 28 (rf OR-cont), Chunk 29 (rf RD), Chunk 39 (OR.ci_plot with # of proxy), Chunk 40 (summary RD all methods), Chunk 41 (print RD result table), chunk 42 (print OR.ci_plot with # of proxy)

---
title: "Step1_LoadData"
author: "Yang Lei"
date: "2024-07-11"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(autoCovariateSelection)
require(dplyr)
require(cobalt)
require(WeightIt)
library(tidyverse)
library(ggplot2)
library(lmtest)
library(MASS)
library(Boruta)
library(GA)
library(mclust)
library(penalizedSVM)
library(xgboost)
library(caret)
require(autoCovariateSelection)
library(randomForest)
```

**Codechuck 0. Loading data -- Generating Recurrence Covariates**

_Step 0: Analytic data, Step 1: Proxy sources, Step 2: Emperical, Step 3: Recurrence_
```{r loading-data, warning=FALSE, cache=TRUE}
# ------------------- Select proxy based on the results from out2
# ------------------- change the code after Step 3: Recurrence
load("data/analytic3cycles.RData")

analytic <- data.complete
idx <- analytic$id
outcome <- as.numeric(analytic$diabetes == "Yes") 
exposure <- as.numeric(analytic$obese == "Yes")

domain <- "dx"
analytic.dfx <- as.data.frame(cbind(idx, exposure, outcome, domain))
dat.proxy.long <- subset(dat.proxy.long, 
                         icd10 != "E66") # Overweight and obesity
dat.proxy.long <- subset(dat.proxy.long, 
                         icd10 != "O24") # Gestational diabetes mellitus
dat.proxy.long <- subset(dat.proxy.long, 
                         icd10 != "E10") # Type 1 diabetes mellitus
dat.proxy.long <- subset(dat.proxy.long, 
                         icd10 != "E11") # Type 2 diabetes mellitus
proxy.var.long <- dat.proxy.long
proxy.var.long$idx <- proxy.var.long$id
proxy.var.long$id <- NULL
dfx <- merge(analytic.dfx, proxy.var.long, by = "idx")
# ------------------- now the proxy data was combined to the A and Y

# basetable <- dfx %>% select(idx, exposure, outcome) %>% distinct()
basetable <- dfx[, c(1:3)] %>% distinct()
patientIds <- basetable$idx
length(patientIds)
# ------------------- number of observations

save(dfx, patientIds, basetable, file = "data/dfx.RData")
load(file = "data/dfx.RData")

step1 <- get_candidate_covariates(df = dfx,  
                                  domainVarname = "domain",
                                  eventCodeVarname = "icd10", 
                                  patientIdVarname = "idx",
                                  patientIdVector = patientIds,
                                  n = 200, 
                                  min_num_patients = 20)
out1 <- step1$covars_data
# ------------------- top 200 prevalent icd10 with min 20 patients

step2 <- get_recurrence_covariates(df = out1, 
                                   patientIdVarname = "idx",
                                   eventCodeVarname = "icd10", 
                                   patientIdVector = patientIds)
out2 <- step2$recurrence_data

exposure <- "obese"
outcome <- "diabetes" 
investigator.specified.covariates <- 
  c(# Demographic
    "age.cat", "sex", "education", "race", 
    "marital", "income", "born", "year",

    # health history related variables/access
    "diabetes.family.history", "medical.access",

    # behavioral
    "smoking", "diet.healthy", "physical.activity", "sleep",

    # Laboratory 
    "uric.acid", "protein.total", "bilirubin.total", "phosphorus",
    "sodium", "potassium", "globulin", "calcium.total", 
    "systolicBP", "diastolicBP", "high.cholesterol"
  )
```

**Method 1: Bross Formula**
```{r bross OR, warning=FALSE, cache=TRUE}
start_time <- Sys.time()

out3 <- get_prioritised_covariates(df = out2,
                                   patientIdVarname = "idx", 
                                   exposureVector = basetable$exposure,
                                   outcomeVector = basetable$outcome,
                                   patientIdVector = patientIds, 
                                   k = 100)
hdps.dim <- out3$autoselected_covariate_df

load("data/analytic3cycles.RData")
hdps.dim$id <- hdps.dim$idx
hdps.dim$idx <- NULL
hdps.data <- merge(data.complete[,c("id",
                                    outcome, 
                                    exposure, 
                                    investigator.specified.covariates)], 
                   hdps.dim, by = "id")

hdps.data$exposure <- as.numeric(I(hdps.data$obese=='Yes'))
hdps.data$outcome <- as.numeric(I(hdps.data$diabetes=='Yes'))

proxy.list.sel <- names(out3$autoselected_covariate_df[,-1])
proxyform <- paste0(proxy.list.sel, collapse = "+")
covform <- paste0(investigator.specified.covariates, collapse = "+")
rhsformula <- paste0(c(covform, proxyform), collapse = "+")
ps.formula <- as.formula(paste0("exposure", "~", rhsformula))

W.out <- weightit(ps.formula, 
                  data = hdps.data, 
                  estimand = "ATE",
                  method = "ps")
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
fit <- glm(out.formula,
           data = hdps.data,
           weights = W.out$weights,
           family= binomial(link = "logit"))
fit.summary <- summary(fit)$coef["exposure",
                                 c("Estimate", 
                                   "Std. Error", 
                                   "Pr(>|z|)")]
fit.ci <- confint(fit, level = 0.95)["exposure", ]
fit.summary_with_ci <- c(fit.summary, fit.ci)
fit.summary_with_ci

fs.bross <- proxy.list.sel
sum_ci.bross <- c(exp(fit.summary_with_ci[1]), fit.summary_with_ci)
names(sum_ci.bross) <- c("Odds Ratio",
                          "Estimate",
                          "Std. Error",
                          "Pr(>|z|)",
                          "2.5 %",
                          "97.5 %")

print("No need to print the number of proxy variables selected, as it is set to be 100.")

print("The Odds Ratio by Bross Formula")
round(sum_ci.bross, 3)
end_time <- Sys.time()
elapsed_time.bross <- as.numeric(difftime(end_time, start_time, units = "secs"))
```

```{r bross RD, warning=FALSE, cache=TRUE}
# Estimating RD and SE
hdps.dim.bross <- out3$autoselected_covariate_df
hdps.dim.bross$id <- hdps.dim.bross$idx
hdps.dim.bross$idx <- NULL
hdps.data.bross <- merge(data.complete[,c("id",
                                    outcome, 
                                    exposure, 
                                    investigator.specified.covariates)], 
                         hdps.dim.bross, 
                         by = "id")
hdps.data.bross$exposure <- as.numeric(I(hdps.data$obese=='Yes'))
hdps.data.bross$outcome <- as.numeric(I(hdps.data$diabetes=='Yes'))

proxyform <- paste0(fs.bross, collapse = "+")
covform <- paste0(investigator.specified.covariates, collapse = "+")
rhsformula <- paste0(c(covform, proxyform), collapse = "+")
ps.formula <- as.formula(paste0("exposure", "~", rhsformula))
out.formula <- as.formula(paste0("outcome","~", "exposure"))

W.out.bross <- weightit(ps.formula, 
                  data = hdps.data.bross, 
                  estimand = "ATE",
                  method = "ps")
fit_gau.bross <- glm(out.formula,
               data= hdps.data.bross,
               weights= W.out.bross$weights,
               family=gaussian(link= "identity"))

summary_rd.bross  <- summary(fit_gau.bross)$coef["exposure",
                                           c("Estimate",
                                             "Std. Error",
                                             "Pr(>|t|)")]
summary_rd.bross[2]<- sqrt(sandwich::sandwich(fit_gau.bross)[2,2])

require(lmtest)
conf.int <- confint(fit_gau.bross, "exposure", level = 0.95,  method = "hc1")
summary_rd.bross  <- c(summary_rd.bross, conf.int)

round(summary_rd.bross, 2)
```

**Method 2: Hybrid Bross Formula & Lasso**
```{r hybrid OR, warning=FALSE, cache=TRUE}
set.seed(42)
start_time <- Sys.time()
proxy.list <- names(out3$autoselected_covariate_df[,-1])
covarsTfull <- c(investigator.specified.covariates, proxy.list)
Y.form <- as.formula(paste0(c("outcome~ exposure", 
                              covarsTfull), collapse = "+") )
covar.mat <- model.matrix(Y.form, data = hdps.data)[,-1]
lasso.fit<-glmnet::cv.glmnet(y = hdps.data$outcome, 
                             x = covar.mat, 
                             type.measure='mse',
                             family="binomial",
                             alpha = 1, 
                             nfolds = 5)
coef.fit<-coef(lasso.fit,s='lambda.min',exact=TRUE)
sel.variables<-row.names(coef.fit)[which(as.numeric(coef.fit)!=0)]
proxy.list.sel.hybrid <- proxy.list[proxy.list %in% sel.variables]
length(proxy.list.sel.hybrid)
fs.hybrid <- proxy.list.sel.hybrid

proxyform <- paste0(proxy.list.sel.hybrid, collapse = "+")
rhsformula <- paste0(c(covform, proxyform), collapse = "+")
ps.formula <- as.formula(paste0("exposure", "~", rhsformula))

W.out <- weightit(ps.formula, 
                  data = hdps.data, 
                  estimand = "ATE",
                  method = "ps")

out.formula <- as.formula(paste0("outcome", "~", "exposure"))
fit <- glm(out.formula,
           data = hdps.data,
           weights = W.out$weights,
           family= binomial(link = "logit"))
fit.summary <- summary(fit)$coef["exposure",
                                 c("Estimate", 
                                   "Std. Error", 
                                   "Pr(>|z|)")]
fit.ci <- confint(fit, level = 0.95)["exposure", ]
fit.summary_with_ci.h <- c(fit.summary, fit.ci)
sum_ci.hybrid <- c(exp(fit.summary_with_ci.h[1]), fit.summary_with_ci.h)

paste("The number of proxy variables selected:", length(proxy.list.sel.hybrid))
print("All proxy variables selected are in common as they are all selected from 
      the proxy variable list by the bross formula")

print("The Odds Ratio by Bross Formula & Lasso")
round(sum_ci.hybrid, 3)
end_time <- Sys.time()
elapsed_time.hybrid <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.hybrid
```

```{r hybrid RD, warning=FALSE, cache=TRUE}
proxyform <- paste0(fs.hybrid, collapse = "+")
covform <- paste0(investigator.specified.covariates, collapse = "+")
rhsformula <- paste0(c(covform, proxyform), collapse = "+")
ps.formula <- as.formula(paste0("exposure", "~", rhsformula))
out.formula <- as.formula(paste0("outcome","~", "exposure"))

W.out.hybrid <- weightit(ps.formula, 
                  data = hdps.data.bross, 
                  estimand = "ATE",
                  method = "ps")
fit_gau.hybrid <- glm(out.formula,
               data= hdps.data.bross,
               weights= W.out.hybrid$weights,
               family=gaussian(link= "identity"))

summary_rd.hybrid  <- summary(fit_gau.hybrid)$coef["exposure",
                                           c("Estimate",
                                             "Std. Error",
                                             "Pr(>|t|)")]
summary_rd.hybrid[2]<- sqrt(sandwich::sandwich(fit_gau.hybrid)[2,2])

require(lmtest)
conf.int <- confint(fit_gau.hybrid, "exposure", level = 0.95,  method = "hc1")
summary_rd.hybrid  <- c(summary_rd.hybrid, conf.int)

round(summary_rd.hybrid, 2)
```

_Before implementing other methods, drop the information from method 1 (Bross formula). Continue with the results from Step 3: Recurrence (table `out2`)._
```{r prepare, warning=FALSE, cache=TRUE}
hdps.dim <- out2

load("data/analytic3cycles.RData")
hdps.dim$id <- hdps.dim$idx
hdps.dim$idx <- NULL
hdps.data <- merge(data.complete[,c("id",
                                    outcome, 
                                    exposure, 
                                    investigator.specified.covariates)], 
                   hdps.dim, by = "id")

hdps.data$exposure <- as.numeric(I(hdps.data$obese=='Yes'))
hdps.data$outcome <- as.numeric(I(hdps.data$diabetes=='Yes'))

proxy.list.sel <- names(out2[,-1])
proxyform <- paste0(proxy.list.sel, collapse = "+")
covform <- paste0(investigator.specified.covariates, collapse = "+")
rhsformula <- paste0(c(covform, proxyform), collapse = "+")
ps.formula <- as.formula(paste0("exposure", "~", rhsformula))

proxy.list <- names(out2[,-1])    # length(proxy.list) :  142
covarsTfull <- c(investigator.specified.covariates, proxy.list)
Y.form <- as.formula(paste0(c("outcome~ exposure", 
                              covarsTfull), collapse = "+") )
```

**Method 3: Pure Lasso**
```{r lasso OR, warning=FALSE, cache=TRUE}
set.seed(42) # different seed -> different length for the selected proxy list
start_time <- Sys.time()
covar.mat <- model.matrix(Y.form, data = hdps.data)[,-1]

lasso.fit <- glmnet::cv.glmnet(y = hdps.data$outcome, 
                               x = covar.mat, 
                               type.measure='mse',
                               family="binomial",
                               alpha = 1, 
                               nfolds = 5)

coef.fit<-coef(lasso.fit,s='lambda.min',exact=TRUE)
sel.variables<-row.names(coef.fit)[which(as.numeric(coef.fit)!=0)]
proxy.list.sel.ml <- proxy.list[proxy.list %in% sel.variables]
length(proxy.list.sel.ml)
fs.lasso <- proxy.list.sel.ml
proxy.lasso <- paste0(fs.lasso, collapse = "+")
form.lasso <- paste0(c(covform, proxy.lasso), collapse = "+")
formula.lasso <- as.formula(paste0("exposure", "~", form.lasso))

W.out.lasso <- weightit(formula.lasso, 
                        data = hdps.data, 
                        estimand = "ATE",
                        method = "ps")

hdps.data$ps.l <- W.out.lasso$ps
hdps.data$w.l <- W.out.lasso$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.lasso <- glm(out.formula,
                 data = hdps.data,
                 weights = W.out.lasso$weights,
                 family= binomial(link = "logit"))
summary.lasso <- summary(fit.lasso)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.lasso <- confint(fit.lasso, level = 0.95)["exposure", ]
sum_ci.lasso <- c(exp(summary.lasso[1]), summary.lasso, ci.lasso)
names(sum_ci.lasso) <- c("Odds Ratio",
                         "Estimate",
                         "Std. Error",
                         "Pr(>|z|)",
                         "2.5 %",
                         "97.5 %")

# Q1: How many proxy variables are in common?
# length(fs.lasso)  # 60 features selected by lasso
# length(fs.lasso[fs.lasso %in% fs.bross])  # 47 of 60 also selected by bross
print("How many proxy variables are in common?")
paste("There are", length(fs.lasso), "features selected by lasso")
paste(length(fs.lasso[fs.lasso %in% fs.bross]), "of the", length(fs.lasso), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.lasso[!(fs.lasso %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Method Pure Lasso")
sum_ci.lasso
end_time <- Sys.time()
elapsed_time.lasso <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.lasso
```

```{r lasso RD, warning=FALSE, cache=TRUE}
fit_gau.lasso <- glm(out.formula,
                     data = hdps.data,
                     weights = W.out.lasso$weights,
                     family = gaussian(link= "identity"))

summary_rd.lasso  <- summary(fit_gau.lasso)$coef["exposure",
                                                 c("Estimate",
                                                   "Std. Error",
                                                   "Pr(>|t|)")]
summary_rd.lasso[2]<- sqrt(sandwich::sandwich(fit_gau.lasso)[2,2])

require(lmtest)
conf.int <- confint(fit_gau.lasso, "exposure", level = 0.95,  method = "hc1")
summary_rd.lasso  <- c(summary_rd.lasso, conf.int)

round(summary_rd.lasso, 2)
```

**Method 4: Elastic Net**
```{r enet OR, warning=FALSE, cache=TRUE}
set.seed(42)
start_time <- Sys.time()
covar.mat <- model.matrix(Y.form, data = hdps.data)[,-1]

elas.fit <- glmnet::cv.glmnet(y = hdps.data$outcome, 
                             x = covar.mat, 
                             type.measure='mse',
                             family="binomial",
                             alpha = 0.5, 
                             nfolds = 5)

coef.fit<-coef(elas.fit, s='lambda.min', exact=TRUE)
sel.variables<-row.names(coef.fit)[which(as.numeric(coef.fit)!=0)]
proxy.list.sel.elastic <- proxy.list[proxy.list %in% sel.variables]
length(proxy.list.sel.elastic)
fs.elastic <- proxy.list.sel.elastic

proxy.elastic <- paste0(fs.elastic, collapse = "+")
form.elastic <- paste0(c(covform, proxy.elastic), collapse = "+")
formula.elastic <- as.formula(paste0("exposure", "~", form.elastic))

W.out.elastic <- weightit(formula.elastic, 
                       data = hdps.data, 
                       estimand = "ATE",
                       method = "ps")

hdps.data$ps.e <- W.out.elastic$ps
hdps.data$w.e <- W.out.elastic$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.elastic <- glm(out.formula,
                data = hdps.data,
                weights = W.out.elastic$weights,
                family= binomial(link = "logit"))
summary.elastic <- summary(fit.elastic)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.elastic <- confint(fit.elastic, level = 0.95)["exposure", ]
sum_ci.elastic <- c(exp(summary.elastic[1]), summary.elastic, ci.elastic)
names(sum_ci.elastic) <- c("Odds Ratio",
                         "Estimate",
                         "Std. Error",
                         "Pr(>|z|)",
                         "2.5 %",
                         "97.5 %")

# Q1: How many proxy variables are in common (Elastic Net vs Bross Formula)?
# length(fs.elastic)  # 69 features selected by elastic net
# length(fs.elastic[fs.elastic %in% fs.bross])  # 54 of 69 also selected by bross formula
print("How many proxy variables are in common?")
paste("There are", length(fs.elastic), "features selected by elas")
paste(length(fs.elastic[fs.elastic %in% fs.bross]), "of the", length(fs.elastic), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.elastic[!(fs.elastic %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Elastic Net")
sum_ci.elastic
end_time <- Sys.time()
elapsed_time.enet <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.enet
```

```{r enet RD, warning=FALSE, cache=TRUE}
fit_gau.elastic <- glm(out.formula,
                     data = hdps.data,
                     weights = W.out.elastic$weights,
                     family = gaussian(link= "identity"))

summary_rd.elastic  <- summary(fit_gau.elastic)$coef["exposure",
                                                 c("Estimate",
                                                   "Std. Error",
                                                   "Pr(>|t|)")]
summary_rd.elastic[2]<- sqrt(sandwich::sandwich(fit_gau.elastic)[2,2])

require(lmtest)
conf.int <- confint(fit_gau.elastic, "exposure", level = 0.95,  method = "hc1")
summary_rd.elastic  <- c(summary_rd.elastic, conf.int)

round(summary_rd.elastic, 2)
```

**Method 5: XGBoost**
```{r xgb OR, warning=FALSE, cache=TRUE}
start_time <- Sys.time()

xgb_fit <- xgboost(data = covar.mat, label = hdps.data$outcome, 
                    max.depth = 20, eta = 1, nthread = 2, nrounds = 5, 
                    objective = "binary:logistic")

xgb_imp <- xgb.importance(model = xgb_fit)
xgb.plot.importance(importance_matrix = xgb_imp)
fs.xgboost <- xgb_imp$Feature
fs.xgboost <- proxy.list[proxy.list %in% fs.xgboost]
proxy.xgboost <- paste0(fs.xgboost, collapse = "+")
form.xgboost <- paste0(c(covform, proxy.xgboost), collapse = "+")
formula.xgboost <- as.formula(paste0("exposure", "~", form.xgboost))

W.out.xgboost <- weightit(formula.xgboost, 
                          data = hdps.data, 
                          estimand = "ATE",
                          method = "ps")

hdps.data$ps.xgboost <- W.out.xgboost$ps
hdps.data$w.xgboost <- W.out.xgboost$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.xgboost <- glm(out.formula,
                   data = hdps.data,
                   weights = W.out.xgboost$weights,
                   family= binomial(link = "logit"))
summary.xgboost <- summary(fit.xgboost)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.xgboost <- confint(fit.xgboost, level = 0.95)["exposure", ]
sum_ci.xgboost <- c(exp(summary.xgboost[1]), summary.xgboost, ci.xgboost)
names(sum_ci.xgboost) <- c("Odds Ratio",
                           "Estimate",
                           "Std. Error",
                           "Pr(>|z|)",
                           "2.5 %",
                           "97.5 %")

# Q1: How many proxy variables are in common?
# length(fs.xgboost) 
# length(fs.xgboost[fs.xgboost %in% fs.bross])
print("How many proxy variables are in common?")
paste("There are", length(fs.xgboost), "features selected by xgboost")
paste(length(fs.xgboost[fs.xgboost %in% fs.bross]), "of the", length(fs.xgboost), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.xgboost[!(fs.xgboost %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by XGBoost")
sum_ci.xgboost
end_time <- Sys.time()
elapsed_time.xgb <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.xgb
```

```{r xgb RD, warning=FALSE, cache=TRUE}
fit_gau.xgboost <- glm(out.formula,
                     data = hdps.data,
                     weights = W.out.xgboost$weights,
                     family = gaussian(link= "identity"))

summary_rd.xgboost  <- summary(fit_gau.xgboost)$coef["exposure",
                                                 c("Estimate",
                                                   "Std. Error",
                                                   "Pr(>|t|)")]
summary_rd.xgboost[2]<- sqrt(sandwich::sandwich(fit_gau.xgboost)[2,2])

require(lmtest)
conf.int <- confint(fit_gau.xgboost, "exposure", level = 0.95,  method = "hc1")
summary_rd.xgboost  <- c(summary_rd.xgboost, conf.int)

round(summary_rd.xgboost, 2)
```

**Method 6: GA**

_GA (Genetic Algorithm) has no specific internal model building mechanism and would wrap around whatever model the user specifies_
```{r ga OR, warning=FALSE, cache=TRUE}
start_time <- Sys.time()

gaOpt <- function(vars, IV.train, DV.train) {
	varNames <- colnames(IV.train) #getting names of all variables
	selectedVarNames <- varNames[vars == "1"] # getting names of selected vars from GA
	gaSolutionData <- IV.train[,selectedVarNames] # keeping only those selected vars

	gaDat <- cbind(gaSolutionData,DV.train) # combining selected variables with outcome variable
	gaMod <- glm(DV.train ~ ., family = "binomial", data = gaDat) #build model
	gaProb <- predict(gaMod, IV.train, type = "response") # get probabilities
	gaPred <- ifelse(gaProb >= .8, 1, 0) # get predicted 0s and 1s

	ari <- adjustedRandIndex(gaPred, DV.train)
	return(ari)
}

ga.solution <- ga(fitness = function(vars)
 gaOpt(vars=vars, 
       IV.train=data.frame(covar.mat),
       DV.train=hdps.data$outcome),
 type = "binary", 
 nBits = ncol(covar.mat),
 names = colnames(covar.mat), 
 seed = 42,
 run=5)

allVarNames <- colnames(covar.mat)
sel.variables.GA <- allVarNames[ga.solution@solution[1,]==1]
fs.GA <- proxy.list[proxy.list %in% sel.variables.GA]
proxy.GA <- paste0(fs.GA, collapse = "+")
form.GA <- paste0(c(covform, proxy.GA), collapse = "+")
formula.GA <- as.formula(paste0("exposure", "~", form.GA))

W.out.GA <- weightit(formula.GA, 
                         data = hdps.data, 
                         estimand = "ATE",
                         method = "ps")

hdps.data$ps.ga <- W.out.GA$ps
hdps.data$w.ga <- W.out.GA$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.GA <- glm(out.formula,
                  data = hdps.data,
                  weights = W.out.GA$weights,
                  family= binomial(link = "logit"))
summary.GA <- summary(fit.GA)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.GA <- confint(fit.GA, level = 0.95)["exposure", ]
sum_ci.GA <- c(exp(summary.GA[1]), summary.GA, ci.GA)
names(sum_ci.GA) <- c("Odds Ratio",
                          "Estimate",
                          "Std. Error",
                          "Pr(>|z|)",
                          "2.5 %",
                          "97.5 %")

# Q1: How many proxy variables are in common (GA vs Bross Formula)?
# length(fs.GA)  # 64 features selected by GA
# length(fs.GA[fs.GA %in% fs.bross])  # 44 of 64 also selected by bross formula
print("How many proxy variables are in common?")
paste("There are", length(fs.GA), "features selected by GA")
paste(length(fs.GA[fs.GA %in% fs.bross]), "of the", length(fs.GA), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.GA[!(fs.GA %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by GA:")
sum_ci.GA
end_time <- Sys.time()
elapsed_time.ga <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.ga
```

```{r ga RD, warning=FALSE, cache=TRUE}
fit_gau.GA <- glm(out.formula,
                     data = hdps.data,
                     weights = W.out.GA$weights,
                     family = gaussian(link= "identity"))

summary_rd.GA  <- summary(fit_gau.GA)$coef["exposure",
                                                 c("Estimate",
                                                   "Std. Error",
                                                   "Pr(>|t|)")]
summary_rd.GA[2]<- sqrt(sandwich::sandwich(fit_gau.GA)[2,2])

require(lmtest)
conf.int <- confint(fit_gau.GA, "exposure", level = 0.95,  method = "hc1")
summary_rd.GA  <- c(summary_rd.GA, conf.int)

round(summary_rd.GA, 2)
```

**Method 7: Elastic SCAD SVM**

_Elastic SCAD SVM uses an approach like LASSO and Elastic Net in that it selects variables via the addition of a penalty._

_(Long Running Time)_
```{r svm OR, warning=FALSE, cache=TRUE}
start_time <- Sys.time()
svmTrainOutcome <- hdps.data$outcome
svmTrainOutcome[svmTrainOutcome == 0] <- -1
svmTrainPreds <- data.frame(covar.mat)

bounds <- t(data.frame(log2lambda1=c(-10, 10), log2lambda2=c(-10,10)))
colnames(bounds)<-c("lower", "upper")
svm.model <- svmfs(x=covar.mat,
				   y = svmTrainOutcome,
				   fs.method = "scad+L2",
				   bounds=bounds,
				   grid.search = "interval",
				   inner.val.method = "cv",
				   show = "none",
				   parms.coding = "none",
				   seed=42)
```

```{r svm OR-cont, warning=FALSE, cache=TRUE}
fs.svm <- colnames(covar.mat)[svm.model$model$fit.info$model.list$model$xind]
fs.svm <- proxy.list[proxy.list %in% fs.svm]
proxy.svm <- paste0(fs.svm, collapse = "+")
form.svm <- paste0(c(covform, proxy.svm), collapse = "+")
formula.svm <- as.formula(paste0("exposure", "~", form.svm))

W.out.svm <- weightit(formula.svm, 
                        data = hdps.data, 
                        estimand = "ATE",
                        method = "ps")

hdps.data$ps.svm <- W.out.svm$ps
hdps.data$w.svm <- W.out.svm$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.svm <- glm(out.formula,
                 data = hdps.data,
                 weights = W.out.svm$weights,
                 family= binomial(link = "logit"))
summary.svm <- summary(fit.svm)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.svm <- confint(fit.svm, level = 0.95)["exposure", ]
sum_ci.svm <- c(exp(summary.svm[1]), summary.svm, ci.svm)
names(sum_ci.svm) <- c("Odds Ratio",
                         "Estimate",
                         "Std. Error",
                         "Pr(>|z|)",
                         "2.5 %",
                         "97.5 %")

# Q1: How many proxy variables are in common?
# length(fs.svm) 
# length(fs.svm[fs.svm %in% fs.bross])
print("How many proxy variables are in common?")
paste("There are", length(fs.svm), "features selected by svm")
paste(length(fs.svm[fs.svm %in% fs.bross]), "of the", length(fs.svm), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.svm[!(fs.svm %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Elastic SCAD SVM")
sum_ci.svm
end_time <- Sys.time()
elapsed_time.svm <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.svm
```

```{r svm RD, warning=FALSE, cache=TRUE}
fit_gau.svm <- glm(out.formula,
                     data = hdps.data,
                     weights = W.out.svm$weights,
                     family = gaussian(link= "identity"))

summary_rd.svm  <- summary(fit_gau.svm)$coef["exposure",
                                                 c("Estimate",
                                                   "Std. Error",
                                                   "Pr(>|t|)")]
summary_rd.svm[2]<- sqrt(sandwich::sandwich(fit_gau.svm)[2,2])

require(lmtest)
conf.int <- confint(fit_gau.svm, "exposure", level = 0.95,  method = "hc1")
summary_rd.svm  <- c(summary_rd.svm, conf.int)

round(summary_rd.svm, 2)
```

**Method 8: Boruta Selection**

_Boruta Selection would give the feature importance ranking based on the Random Forests algorithm which are able to capture some non-linear trends but not easy to interpret._

_(Long Running Time)_
```{r Boruta OR, warning=FALSE, cache=TRUE}
set.seed(42)
start_time <- Sys.time()

boruta_output <- Boruta(Y.form, data = hdps.data, doTrace = 1)

sel.variables.bor <- names(boruta_output$finalDecision)[which(boruta_output$finalDecision != "Rejected")]
fs.boruta <- proxy.list[proxy.list %in% sel.variables.bor]
proxy.boruta <- paste0(fs.boruta, collapse = "+")
form.boruta <- paste0(c(covform, proxy.boruta), collapse = "+")
formula.boruta <- as.formula(paste0("exposure", "~", form.boruta))

W.out.boruta <- weightit(formula.boruta, 
                         data = hdps.data, 
                         estimand = "ATE",
                         method = "ps")

hdps.data$ps.b <- W.out.boruta$ps
hdps.data$w.b <- W.out.boruta$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.boruta <- glm(out.formula,
                  data = hdps.data,
                  weights = W.out.boruta$weights,
                  family= binomial(link = "logit"))
summary.boruta <- summary(fit.boruta)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.boruta <- confint(fit.boruta, level = 0.95)["exposure", ]
sum_ci.boruta <- c(exp(summary.boruta[1]), summary.boruta, ci.boruta)
names(sum_ci.boruta) <- c("Odds Ratio",
                          "Estimate",
                          "Std. Error",
                          "Pr(>|z|)",
                          "2.5 %",
                          "97.5 %")

# Q1: How many proxy variables are in common (Boruta vs Bross Formula)?
# length(fs.boruta)  # 28 features selected by boruta selection
# length(fs.boruta[fs.boruta %in% fs.bross])  # 24 of 28 also selected by bross formula
print("How many proxy variables are in common?")
paste("There are", length(fs.boruta), "features selected by boruta")
paste(length(fs.boruta[fs.boruta %in% fs.bross]), "of the", length(fs.boruta), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.boruta[!(fs.boruta %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Boruta Selection")
sum_ci.boruta
end_time <- Sys.time()
elapsed_time.boruta <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.boruta
```

```{r Boruta RD, warning=FALSE, cache=TRUE}
fit_gau.boruta <- glm(out.formula,
                     data = hdps.data,
                     weights = W.out.boruta$weights,
                     family = gaussian(link= "identity"))

summary_rd.boruta  <- summary(fit_gau.boruta)$coef["exposure",
                                                 c("Estimate",
                                                   "Std. Error",
                                                   "Pr(>|t|)")]
summary_rd.boruta[2]<- sqrt(sandwich::sandwich(fit_gau.boruta)[2,2])

require(lmtest)
conf.int <- confint(fit_gau.boruta, "exposure", level = 0.95,  method = "hc1")
summary_rd.boruta  <- c(summary_rd.boruta, conf.int)

round(summary_rd.boruta, 2)
```

**Method 9.1: Stepwise Regression (Forward)**
```{r fs OR, warning=FALSE, cache=TRUE}
set.seed(42)
start_time <- Sys.time()

initial_formula <- as.formula(paste0("outcome~exposure+",
                                     covform,
                                     collapse = "+"))
full_formula <- as.formula(paste0("outcome~exposure+",
                                  rhsformula,
                                  collapse = "+"))

initial_model <- glm(initial_formula, data = hdps.data, family = binomial)
full_model <- glm(full_formula, data = hdps.data, family = binomial)
stepwise_forward <- stepAIC(initial_model, 
                            scope = list(lower = initial_model, 
                                         upper = full_model), 
                            direction = "forward")

fs.forward <- all.vars(formula(stepwise_forward))[-1]
fs.forward <- proxy.list[proxy.list %in% fs.forward]
proxy.forward <- paste0(fs.forward, collapse = "+")
form.forward <- paste0(c(covform, proxy.forward), collapse = "+")
formula.forward <- as.formula(paste0("exposure", "~", form.forward))

W.out.forward <- weightit(formula.forward, 
                          data = hdps.data, 
                          estimand = "ATE",
                          method = "ps")

hdps.data$ps.forward <- W.out.forward$ps
hdps.data$w.forward <- W.out.forward$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.forward <- glm(out.formula,
                   data = hdps.data,
                   weights = W.out.forward$weights,
                   family= binomial(link = "logit"))
summary.forward <- summary(fit.forward)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.forward <- confint(fit.forward, level = 0.95)["exposure", ]
sum_ci.forward <- c(exp(summary.forward[1]), summary.forward, ci.forward)
names(sum_ci.forward) <- c("Odds Ratio",
                           "Estimate",
                           "Std. Error",
                           "Pr(>|z|)",
                           "2.5 %",
                           "97.5 %")

# Q1: How many proxy variables are in common?
# length(fs.forward) 
# length(fs.forward[fs.forward %in% fs.bross])
print("How many proxy variables are in common?")
paste("There are", length(fs.forward), "features selected by forward")
paste(length(fs.forward[fs.forward %in% fs.bross]), "of the", length(fs.forward), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.forward[!(fs.forward %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Stepwise Forward Selection")
sum_ci.forward
end_time <- Sys.time()
elapsed_time.fs <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.fs
```

```{r fs RD, warning=FALSE, cache=TRUE}
fit_gau.forward <- glm(out.formula,
                     data = hdps.data,
                     weights = W.out.forward$weights,
                     family = gaussian(link= "identity"))

summary_rd.forward  <- summary(fit_gau.forward)$coef["exposure",
                                                 c("Estimate",
                                                   "Std. Error",
                                                   "Pr(>|t|)")]
summary_rd.forward[2]<- sqrt(sandwich::sandwich(fit_gau.forward)[2,2])

require(lmtest)
conf.int <- confint(fit_gau.forward, "exposure", level = 0.95,  method = "hc1")
summary_rd.forward  <- c(summary_rd.forward, conf.int)

round(summary_rd.forward, 2)
```

**Method 9.2: Stepwise Regression (Backward)**
```{r bs OR, warning=FALSE, cache=TRUE}
start_time <- Sys.time()

stepwise_backward <- stepAIC(full_model, direction = "backward")

fs.backward <- all.vars(formula(stepwise_backward))[-1]
fs.backward <- proxy.list[proxy.list %in% fs.backward]
fs.backward <- fs.backward[fs.backward != "exposure"]
proxy.backward <- paste0(fs.backward, collapse = "+")
form.backward <- paste0(c(covform, proxy.backward), collapse = "+")
formula.backward <- as.formula(paste0("exposure", "~", form.backward))

W.out.backward <- weightit(formula.backward, 
                          data = hdps.data, 
                          estimand = "ATE",
                          method = "ps")

hdps.data$ps.backward <- W.out.backward$ps
hdps.data$w.backward <- W.out.backward$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.backward <- glm(out.formula,
                   data = hdps.data,
                   weights = W.out.backward$weights,
                   family= binomial(link = "logit"))
summary.backward <- summary(fit.backward)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.backward <- confint(fit.backward, level = 0.95)["exposure", ]
sum_ci.backward <- c(exp(summary.backward[1]), summary.backward, ci.backward)
names(sum_ci.backward) <- c("Odds Ratio",
                           "Estimate",
                           "Std. Error",
                           "Pr(>|z|)",
                           "2.5 %",
                           "97.5 %")

# Q1: How many proxy variables are in common?
# length(fs.backward) 
# length(fs.backward[fs.backward %in% fs.bross])
print("How many proxy variables are in common?")
paste("There are", length(fs.backward), "features selected by backward")
paste(length(fs.backward[fs.backward %in% fs.bross]), "of the", length(fs.backward), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.backward[!(fs.backward %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Stepwise Backward Selection")
sum_ci.backward
end_time <- Sys.time()
elapsed_time.bs <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.bs
```

```{r bs RD, warning=FALSE, cache=TRUE}
fit_gau.backward <- glm(out.formula,
                     data = hdps.data,
                     weights = W.out.backward$weights,
                     family = gaussian(link= "identity"))

summary_rd.backward  <- summary(fit_gau.backward)$coef["exposure",
                                                 c("Estimate",
                                                   "Std. Error",
                                                   "Pr(>|t|)")]
summary_rd.backward[2]<- sqrt(sandwich::sandwich(fit_gau.backward)[2,2])

require(lmtest)
conf.int <- confint(fit_gau.backward, "exposure", level = 0.95,  method = "hc1")
summary_rd.backward  <- c(summary_rd.backward, conf.int)

round(summary_rd.backward, 2)
```

**Method 9.3: Stepwise Regression (Both)**
```{r sw OR, warning=FALSE, cache=TRUE}
start_time <- Sys.time()

stepwise_both <- stepAIC(initial_model, 
                            scope = list(lower = initial_model, 
                                         upper = full_model), 
                            direction = "both")

fs.both <- all.vars(formula(stepwise_both))[-1]
fs.both <- proxy.list[proxy.list %in% fs.both]
proxy.both <- paste0(fs.both, collapse = "+")
form.both <- paste0(c(covform, proxy.both), collapse = "+")
formula.both <- as.formula(paste0("exposure", "~", form.both))

W.out.both <- weightit(formula.both, 
                          data = hdps.data, 
                          estimand = "ATE",
                          method = "ps")

hdps.data$ps.both <- W.out.both$ps
hdps.data$w.both <- W.out.both$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
out.formula
fit.both <- glm(out.formula,
                   data = hdps.data,
                   weights = W.out.both$weights,
                   family= binomial(link = "logit"))
summary.both <- summary(fit.both)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci.both <- confint(fit.both, level = 0.95)["exposure", ]
sum_ci.both <- c(exp(summary.both[1]), summary.both, ci.both)
names(sum_ci.both) <- c("Odds Ratio",
                           "Estimate",
                           "Std. Error",
                           "Pr(>|z|)",
                           "2.5 %",
                           "97.5 %")

# Q1: How many proxy variables are in common?
# length(fs.both) 
# length(fs.both[fs.both %in% fs.bross])
print("How many proxy variables are in common?")
paste("There are", length(fs.both), "features selected by both")
paste(length(fs.both[fs.both %in% fs.bross]), "of the", length(fs.both), 
      "are also selected by the bross formula method")
# See which variables are not in common:
# fs.both[!(fs.both %in% fs.bross)]

# Q2: Odds ratio
print("The Odds Ratio by Stepwise Both Selection")
sum_ci.both
end_time <- Sys.time()
elapsed_time.sw <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.sw
```

```{r sw RD, warning=FALSE, cache=TRUE}
fit_gau.both <- glm(out.formula,
                     data = hdps.data,
                     weights = W.out.both$weights,
                     family = gaussian(link= "identity"))

summary_rd.both  <- summary(fit_gau.both)$coef["exposure",
                                                 c("Estimate",
                                                   "Std. Error",
                                                   "Pr(>|t|)")]
summary_rd.both[2]<- sqrt(sandwich::sandwich(fit_gau.both)[2,2])

require(lmtest)
conf.int <- confint(fit_gau.both, "exposure", level = 0.95,  method = "hc1")
summary_rd.both  <- c(summary_rd.both, conf.int)

round(summary_rd.both, 2)
```

```{r rf OR, warning=FALSE, cache=TRUE}
# Random Forest Feature Selection
start_time <- Sys.time()

rf_model <- randomForest(outcome ~ ., data = hdps.data, importance = TRUE)

# Get important features
rf_importance <- importance(rf_model)
important_features_rf <- rownames(rf_importance)[order(rf_importance[, 1], decreasing = TRUE)]
```

```{r rf OR-cont, warning=FALSE, cache=TRUE}
# Select features
fs_rf <- important_features_rf[important_features_rf %in% proxy.list][1:100]
proxy_rf <- paste0(fs_rf, collapse = "+")
form_rf <- paste0(c(covform, proxy_rf), collapse = "+")
formula_rf <- as.formula(paste0("exposure", "~", form_rf))

W.out_rf <- weightit(formula_rf, 
                     data = hdps.data, 
                     estimand = "ATE",
                     method = "ps")

hdps.data$ps_rf <- W.out_rf$ps
hdps.data$w_rf <- W.out_rf$weights
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
fit_rf <- glm(out.formula,
              data = hdps.data,
              weights = W.out_rf$weights,
              family= binomial(link = "logit"))
summary_rf <- summary(fit_rf)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
ci_rf <- confint(fit_rf, level = 0.95)["exposure", ]
sum_ci_rf <- c(exp(summary_rf[1]), summary_rf, ci_rf)
names(sum_ci_rf) <- c("Odds Ratio",
                      "Estimate",
                      "Std. Error",
                      "Pr(>|z|)",
                      "2.5 %",
                      "97.5 %")

end_time <- Sys.time()
elapsed_time_rf <- as.numeric(difftime(end_time, start_time, units = "secs"))

print("How many proxy variables are in common?")
paste("There are", length(fs_rf), "features selected by randomForest")
paste(length(fs_rf[fs_rf %in% fs.bross]), "of the", length(fs_rf), 
      "are also selected by the bross formula method")

print("The Odds Ratio by Random Forest")
sum_ci_rf
print("Elapsed time for Random Forest:")
elapsed_time_rf
```

```{r rf RD, warning=FALSE, cache=TRUE}
fit_gau.rf <- glm(out.formula,
                     data = hdps.data,
                     weights = W.out_rf$weights,
                     family = gaussian(link= "identity"))

summary_rd.rf  <- summary(fit_gau.rf)$coef["exposure",
                                                 c("Estimate",
                                                   "Std. Error",
                                                   "Pr(>|t|)")]
summary_rd.rf[2]<- sqrt(sandwich::sandwich(fit_gau.rf)[2,2])

require(lmtest)
conf.int <- confint(fit_gau.rf, "exposure", level = 0.95,  method = "hc1")
summary_rd.rf  <- c(summary_rd.rf, conf.int)

round(summary_rd.rf, 2)
```

Naming consistence:
```{r}
sum_ci.rf <- sum_ci_rf
fs.rf <- fs_rf
```

Comparing the Results:
```{r compare, warning=FALSE, cache=TRUE}
sum_ci.all <- rbind(sum_ci.bross, sum_ci.lasso, sum_ci.hybrid, sum_ci.elastic, 
                    sum_ci.svm, sum_ci.GA, 
                    sum_ci.xgboost, sum_ci.rf, sum_ci.boruta,
                    sum_ci.forward, sum_ci.backward, sum_ci.both)

OR.data <- data.frame(method = c("bross", "lasso", "hybrid", "elastic", 
                                 "SVM", "GA",
                                 "xgboost", "rf", "boruta", 
                                 "forward", "backward", "both"),
                      OR = exp(sum_ci.all[,2]),
                      lower = exp(sum_ci.all[,5]),
                      upper = exp(sum_ci.all[,6]))

OR.data$method <- factor(OR.data$method, 
                         levels = c("both", "backward", "forward",
                                    "boruta", "rf", "xgboost", 
                                    "GA", "SVM", 
                                    "elastic", "hybrid", "lasso", "bross"))

OR.ci_plot <- ggplot(OR.data, aes(y = method, x = OR)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.2) + 
  labs(title = "Odds Ratios with 95% Confidence Intervals",
       x = "Odds Ratio",
       y = "Method")

fs.all <- data.frame(numFeature = c(length(fs.bross), 
                                    length(fs.lasso),
                                    length(fs.hybrid),
                                    length(fs.elastic), 
                                    length(fs.svm),
                                    length(fs.GA),
                                    length(fs.xgboost),
                                    length(fs.rf),
                                    length(fs.boruta),
                                    length(fs.forward),
                                    length(fs.backward),
                                    length(fs.both)),
                     numInCommon = c(NA,
                                     length(fs.lasso[fs.lasso %in% fs.bross]),
                                     length(fs.hybrid[fs.hybrid %in% fs.bross]),
                                     length(fs.elastic[fs.elastic %in% fs.bross]),
                                     length(fs.svm[fs.svm %in% fs.bross]),
                                     length(fs.GA[fs.GA %in% fs.bross]),
                                     length(fs.xgboost[fs.xgboost %in% fs.bross]),
                                     length(fs.rf[fs.rf %in% fs.bross]),
                                     length(fs.boruta[fs.boruta %in% fs.bross]),
                                     length(fs.forward[fs.forward %in% fs.bross]),
                                     length(fs.backward[fs.backward %in% fs.bross]),
                                     length(fs.both[fs.both %in% fs.bross])))
rownames(fs.all) <-  c("bross", "lasso", "hybrid", "elastic", 
                       "SVM", "GA",
                       "xgboost", "rf", "boruta", 
                       "forward", "backward", "both")

fs.all$pctCommon <- fs.all$numInCommon / fs.all$numFeature

fs.commonOneToOne <- data.frame(ini = rep(NA, 12))

for (i in 1:12) {
  col_name <- OR.data$method[i]
  if (col_name == "SVM") {
    col_name <- "svm"
  }
  na.i <- 1
  na <- c()
  while (na.i < i) {
    na <- c(na, NA)
    na.i <- na.i + 1
  }
  num_common <- c(length(fs.bross[fs.bross %in% get(paste0("fs.",col_name))]),
                  length(fs.lasso[fs.lasso %in% get(paste0("fs.",col_name))]),
                  length(fs.hybrid[fs.hybrid %in% get(paste0("fs.",col_name))]),
                  length(fs.elastic[fs.elastic %in% get(paste0("fs.",col_name))]),
                  length(fs.svm[fs.svm %in% get(paste0("fs.",col_name))]),
                  length(fs.GA[fs.GA %in% get(paste0("fs.",col_name))]),
                  length(fs.xgboost[fs.xgboost %in% get(paste0("fs.",col_name))]),
                  length(fs.rf[fs.rf %in% get(paste0("fs.",col_name))]),
                  length(fs.boruta[fs.boruta %in% get(paste0("fs.",col_name))]),
                  length(fs.forward[fs.forward %in% get(paste0("fs.",col_name))]),
                  length(fs.backward[fs.backward %in% get(paste0("fs.",col_name))]),
                  length(fs.both[fs.both %in% get(paste0("fs.",col_name))]))
  fs.commonOneToOne[i] <- c(na, num_common[i:12])
}

colnames(fs.commonOneToOne) <- rownames(fs.all)
rownames(fs.commonOneToOne) <- rownames(fs.all)
```

Print the comparing plot & table:
```{r}
elapsed_times <- list()
elapsed_times[["bross"]] <- elapsed_time.bross
elapsed_times[["hybrid"]] <- elapsed_time.hybrid
elapsed_times[["lasso"]] <- elapsed_time.lasso
elapsed_times[["enet"]] <- elapsed_time.enet
elapsed_times[["xgb"]] <- elapsed_time.xgb
elapsed_times[["ga"]] <- elapsed_time.ga
elapsed_times[["svm"]] <- elapsed_time.svm
elapsed_times[["boruta"]] <- elapsed_time.boruta
elapsed_times[["fs"]] <- elapsed_time.fs
elapsed_times[["bs"]] <- elapsed_time.bs
elapsed_times[["sw"]] <- elapsed_time.sw
print(elapsed_times)
print("Comparison of odds ratios and the confidence intervals by different proxy selection methods:")
OR.ci_plot
```

```{r}
print("Comparison of the number & percentage of proxy variables selected by different methods in common with that by the bross formula:")
fs.all
```

```{r}
print("One-to-one comparison of the number of proxy variables selected by different methods")
fs.commonOneToOne
```

```{r rfe, warning=FALSE, cache=TRUE}
# RFE Feature Selection
start_time <- Sys.time()
control <- rfeControl(functions = rfFuncs, method = "cv", number = 10)
results <- rfe(hdps.data[, -which(names(hdps.data) %in% c("outcome", "exposure"))], hdps.data$outcome, sizes = c(1:10), rfeControl = control)

# Get important features
important_features_rfe <- predictors(results)
```

```{r}
important_features_rfe
```

Why does this codechunk not work?
No proxy is selected by rfe. it only selects diabetes.
ERROR: the formula in wrong format (ending with "+")
" e ~ a+b+c+ "

```{r rfe2, warning=FALSE, cache=TRUE}
# Select features
# fs_rfe <- important_features_rfe[important_features_rfe %in% proxy.list]
# proxy_rfe <- paste0(fs_rfe, collapse = "+")
# form_rfe <- paste0(c(covform, proxy_rfe), collapse = "+")
# formula_rfe <- as.formula(paste0("exposure", "~", form_rfe))
# 
# W.out_rfe <- weightit(formula_rfe, 
#                       data = hdps.data, 
#                       estimand = "ATE",
#                       method = "ps")
# 
# hdps.data$ps_rfe <- W.out_rfe$ps
# hdps.data$w_rfe <- W.out_rfe$weights
# out.formula <- as.formula(paste0("outcome", "~", "exposure"))
# fit_rfe <- glm(out.formula,
#                data = hdps.data,
#                weights = W.out_rfe$weights,
#                family= binomial(link = "logit"))
# summary_rfe <- summary(fit_rfe)$coef["exposure", c("Estimate", "Std. Error", "Pr(>|z|)")]
# ci_rfe <- confint(fit_rfe, level = 0.95)["exposure", ]
# sum_ci_rfe <- c(exp(summary_rfe[1]), summary_rfe, ci_rfe)
# names(sum_ci_rfe) <- c("Odds Ratio",
#                        "Estimate",
#                        "Std. Error",
#                        "Pr(>|z|)",
#                        "2.5 %",
#                        "97.5 %")
# 
# end_time <- Sys.time()
# elapsed_time_rfe <- as.numeric(difftime(end_time, start_time, units = "secs"))
# 
# print("How many proxy variables are in common?")
# paste("There are", length(fs_rfe), "features selected by RFE")
# paste(length(fs_rfe[fs_rfe %in% fs.bross]), "of the", length(fs_rfe), 
#       "are also selected by the bross formula method")
# 
# print("The Odds Ratio by RFE")
# sum_ci_rfe
# print("Elapsed time for RFE:")
# elapsed_time_rfe
```

```{r save}
# Save the entire workspace to a file named "workspace.RData"
save.image(file = "data/workspace.RData")
```

```{r OR.ci_plot with # of proxy}

OR.ci_plot_new <- ggplot(OR.data, aes(y = method, x = OR)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.2) + 
  geom_text(aes(label = fs.all[,1], x = OR, y = method),
            vjust = -0.8,  # Adjust vertical position
            hjust = -1, # Adjust horizontal position
            size = 3, 
            color = "dark green") +
  labs(title = "Odds Ratios with 95% Confidence Intervals",
       x = "Odds Ratio",
       y = "Method")
```

RD comparison:
```{r summary RD all methods}
sum.rd_all <- rbind(summary_rd.bross,
                    summary_rd.hybrid,
                    summary_rd.lasso,
                    summary_rd.elastic,
                    summary_rd.svm,
                    summary_rd.GA,
                    summary_rd.xgboost,
                    summary_rd.rf,
                    summary_rd.boruta,
                    summary_rd.forward,
                    summary_rd.backward,
                    summary_rd.both)

rownames(sum.rd_all) <- c("bross", "hybrid", "lasso", "elastic", 
                          "SVM", "GA",
                          "xgboost", "rf", "boruta", 
                          "forward", "backward", "both")
```

```{r save RD}
# Save the entire workspace to a file named "workspaceNew.RData"
save.image(file = "data/workspaceNew.RData")
```

```{r print RD result table}
round(sum.rd_all, 3)
```

```{r print OR.ci_plot with # of proxy}
OR.ci_plot_new
```