### 2024-08-01 sim_6.ga/simResults_ga.Rmd
sim results svm, ga

---
title: "simResults_ga"
author: "Yang Lei"
date: "2024-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(autoCovariateSelection)
require(dplyr)
require(cobalt)
require(WeightIt)
library(tidyverse)
library(ggplot2)
library(lmtest)
library(MASS)
library(Boruta)
library(GA)
library(mclust)
library(penalizedSVM)
library(xgboost)
library(caret)
require(autoCovariateSelection)
library(randomForest)
require(lmtest)
```

```{r load basic information, warning=FALSE, cache=TRUE}
# global information
exposure <- "obese"
outcome <- "diabetes" 
investigator.specified.covariates <- 
  c(# Demographic
    "age.cat", "sex", "education", "race", 
    "marital", "income", "born", "year",

    # health history related variables/access
    "diabetes.family.history", "medical.access",

    # behavioral
    "smoking", "diet.healthy", "physical.activity", "sleep",

    # Laboratory 
    "uric.acid", "protein.total", "bilirubin.total", "phosphorus",
    "sodium", "potassium", "globulin", "calcium.total", 
    "systolicBP", "diastolicBP", "high.cholesterol"
  )
covform <- paste0(investigator.specified.covariates, collapse = "+")
out.formula <- as.formula(paste0("outcome", "~", "exposure"))

url <- paste0("https://raw.githubusercontent.com/ehsanx/PlasmodeSim/main/data/hdData/data/scenario/data_1.rds")
data <- readRDS(url(url))
proxy.list <- names(data[, c(grep("^rec", names(data), value = TRUE))])
covarsTfull <- c(investigator.specified.covariates, proxy.list)
Y.form <- as.formula(paste0(c("outcome~ exposure", 
                              covarsTfull), collapse = "+") )

# function for GA
gaOpt <- function(vars, IV.train, DV.train) {
	varNames <- colnames(IV.train) #getting names of all variables
	selectedVarNames <- varNames[vars == "1"] # getting names of selected vars from GA
	gaSolutionData <- IV.train[,selectedVarNames] # keeping only those selected vars

	gaDat <- cbind(gaSolutionData,DV.train) # combining selected variables with outcome variable
	gaMod <- glm(DV.train ~ ., family = "binomial", data = gaDat) #build model
	gaProb <- predict(gaMod, IV.train, type = "response") # get probabilities
	gaPred <- ifelse(gaProb >= .8, 1, 0) # get predicted 0s and 1s

	ari <- adjustedRandIndex(gaPred, DV.train)
	return(ari)
}
```

```{r initialization before for-loop, warning=FALSE, cache=TRUE}
# 'a' ranges from 1 to 1000 to change the number of datasets loaded in each for-loop
a <- 1000

RD_ga <- data.frame(RD = numeric(),
                       SE = numeric())
```

```{r scenario: for-loop generating RD & SE results for ga, warning=FALSE, cache=TRUE}
set.seed(42)

# Folder "scenario"
for (i in 1:a) {
  url <- paste0("https://raw.githubusercontent.com/ehsanx/PlasmodeSim/main/data/hdData/data/scenario/data_", 
                i, 
                ".rds")
  data <- tryCatch({
    readRDS(url(url))
  }, error = function(e) {
    closeAllConnections()
    readRDS(url(url))
  })

  # found some id != idx
  data$id <- data$idx
  covar.mat <- model.matrix(Y.form, data = data)[,-1]

  ga.solution <- ga(fitness = function(vars)
    gaOpt(vars = vars, 
          IV.train = data.frame(covar.mat),
          DV.train = data$outcome),
    type = "binary", 
    nBits = ncol(covar.mat),
    names = colnames(covar.mat), 
    seed = 42,
    run=5)

  allVarNames <- colnames(covar.mat)
  sel.variables <- allVarNames[ga.solution@solution[1,]==1]
  proxy_ga <- proxy.list[proxy.list %in% sel.variables]
  proxyform <- paste0(proxy_ga, collapse = "+")
  rhsform <- paste0(c(covform, proxyform), collapse = "+")
  ps.formula <- as.formula(paste0("exposure", "~", rhsform))

  W.out_ga <- weightit(ps.formula, 
                       data = data, 
                       estimand = "ATE",
                       method = "ps")

  data$ps.l <- W.out_ga$ps
  data$w.l <- W.out_ga$weights

  fit.OR_ga <- glm(out.formula,
                      data = data,
                      weights = W.out_ga$weights,
                      family= binomial(link = "logit"))
  fit.RD_ga <- glm(out.formula,
                       data = data,
                       weights = W.out_ga$weights,
                       family = gaussian(link= "identity"))
  sum.RD_ga <- c(summary(fit.RD_ga)$coef["exposure", c("Estimate")], sqrt(sandwich::sandwich(fit.RD_ga)[2,2]))

  RD_ga[i,] <- sum.RD_ga
  rownames(RD_ga)[i] <- paste0("data.", i, "_ga")

  proxy_ga.i <- paste0("proxy_ga.", i)
  assign(proxy_ga.i, proxy_ga)
  sum.RD_ga.i <- paste0("sum.RD_ga.", i)
  assign(sum.RD_ga.i, sum.RD_ga)
}
```

```{r scenarioER: for-loop generating RD & SE results for ga, warning=FALSE, cache=TRUE}
set.seed(42)

# Folder "scenarioER"
for (i in 1:a) {
  url <- paste0("https://raw.githubusercontent.com/ehsanx/PlasmodeSim/main/data/hdData/data/scenarioER/data_", 
                i, 
                ".rds")
  data <- tryCatch({
    readRDS(url(url))
  }, error = function(e) {
    closeAllConnections()
    readRDS(url(url))
  })

  # found some id != idx
  data$id <- data$idx
  covar.mat <- model.matrix(Y.form, data = data)[,-1]

  ga.solution <- ga(fitness = function(vars)
    gaOpt(vars = vars, 
          IV.train = data.frame(covar.mat),
          DV.train = data$outcome),
    type = "binary", 
    nBits = ncol(covar.mat),
    names = colnames(covar.mat), 
    seed = 42,
    run=5)

  allVarNames <- colnames(covar.mat)
  sel.variables <- allVarNames[ga.solution@solution[1,]==1]
  proxy_ga <- proxy.list[proxy.list %in% sel.variables]
  proxyform <- paste0(proxy_ga, collapse = "+")
  rhsform <- paste0(c(covform, proxyform), collapse = "+")
  ps.formula <- as.formula(paste0("exposure", "~", rhsform))

  W.out_ga <- weightit(ps.formula, 
                       data = data, 
                       estimand = "ATE",
                       method = "ps")

  data$ps.l <- W.out_ga$ps
  data$w.l <- W.out_ga$weights

  fit.OR_ga <- glm(out.formula,
                      data = data,
                      weights = W.out_ga$weights,
                      family= binomial(link = "logit"))
  fit.RD_ga <- glm(out.formula,
                       data = data,
                       weights = W.out_ga$weights,
                       family = gaussian(link= "identity"))
  sum.RD_ga <- c(summary(fit.RD_ga)$coef["exposure", c("Estimate")], sqrt(sandwich::sandwich(fit.RD_ga)[2,2]))

  RD_ga[i+a,] <- sum.RD_ga
  rownames(RD_ga)[i+a] <- paste0("ER.data.", i, "_ga")

  proxy_ga.i <- paste0("proxy_ga.ER.", i)
  assign(proxy_ga.i, proxy_ga)
  sum.RD_ga.i <- paste0("sum.RD_ga.ER.", i)
  assign(sum.RD_ga.i, sum.RD_ga)
}
```

```{r scenarioOR: for-loop generating RD & SE results for ga, warning=FALSE, cache=TRUE}
set.seed(42)

# Folder "scenarioOR"
for (i in 1:a) {
  url <- paste0("https://raw.githubusercontent.com/ehsanx/PlasmodeSim/main/data/hdData/data/scenarioOR/data_", 
                i, 
                ".rds")
  data <- tryCatch({
    readRDS(url(url))
  }, error = function(e) {
    closeAllConnections()
    readRDS(url(url))
  })

  # found some id != idx
  data$id <- data$idx
  covar.mat <- model.matrix(Y.form, data = data)[,-1]

  ga.solution <- ga(fitness = function(vars)
    gaOpt(vars = vars, 
          IV.train = data.frame(covar.mat),
          DV.train = data$outcome),
    type = "binary", 
    nBits = ncol(covar.mat),
    names = colnames(covar.mat), 
    seed = 42,
    run=5)

  allVarNames <- colnames(covar.mat)
  sel.variables <- allVarNames[ga.solution@solution[1,]==1]
  proxy_ga <- proxy.list[proxy.list %in% sel.variables]
  proxyform <- paste0(proxy_ga, collapse = "+")
  rhsform <- paste0(c(covform, proxyform), collapse = "+")
  ps.formula <- as.formula(paste0("exposure", "~", rhsform))

  W.out_ga <- weightit(ps.formula, 
                       data = data, 
                       estimand = "ATE",
                       method = "ps")

  data$ps.l <- W.out_ga$ps
  data$w.l <- W.out_ga$weights

  fit.OR_ga <- glm(out.formula,
                      data = data,
                      weights = W.out_ga$weights,
                      family= binomial(link = "logit"))
  fit.RD_ga <- glm(out.formula,
                       data = data,
                       weights = W.out_ga$weights,
                       family = gaussian(link= "identity"))
  sum.RD_ga <- c(summary(fit.RD_ga)$coef["exposure", c("Estimate")], sqrt(sandwich::sandwich(fit.RD_ga)[2,2]))

  RD_ga[i+2*a,] <- sum.RD_ga
  rownames(RD_ga)[i+2*a] <- paste0("OR.data.", i, "_ga")

  proxy_ga.i <- paste0("proxy_ga.OR.", i)
  assign(proxy_ga.i, proxy_ga)
  sum.RD_ga.i <- paste0("sum.RD_ga.OR.", i)
  assign(sum.RD_ga.i, sum.RD_ga)
}
```

```{r}
RD_ga
```

```{r check intermediate results for each data by ga, warning=FALSE, cache=TRUE}
# check RD & SE by ga for each data, print:
#   sum.RD_ga.n, sum.RD_ga.ER.n, or sum.RD_ga.OR.n
#   where "n" is a number ranges from 1 to 1000
#   e.g.
sum.RD_ga.1
sum.RD_ga.ER.2

# check proxy selected by ga for each data, print:
#   proxy_ga.n, proxy_ga.ER.n, or proxy_ga.OR.n
#   where "n" is a number ranges from 1 to 1000
#   e.g.
proxy_ga.1
proxy_ga.ER.3
```