### 2024-08-01 sim_8.rf/simResults_rf.Rmd
sim results rf

---
title: "simResults_rf"
author: "Yang Lei"
date: "2024-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(autoCovariateSelection)
require(dplyr)
require(cobalt)
require(WeightIt)
library(tidyverse)
library(ggplot2)
library(lmtest)
library(MASS)
library(Boruta)
library(GA)
library(mclust)
library(penalizedSVM)
library(xgboost)
library(caret)
require(autoCovariateSelection)
library(randomForest)
require(lmtest)
```

```{r load basic information, warning=FALSE, cache=TRUE}
# global information
exposure <- "obese"
outcome <- "diabetes" 
investigator.specified.covariates <- 
  c(# Demographic
    "age.cat", "sex", "education", "race", 
    "marital", "income", "born", "year",

    # health history related variables/access
    "diabetes.family.history", "medical.access",

    # behavioral
    "smoking", "diet.healthy", "physical.activity", "sleep",

    # Laboratory 
    "uric.acid", "protein.total", "bilirubin.total", "phosphorus",
    "sodium", "potassium", "globulin", "calcium.total", 
    "systolicBP", "diastolicBP", "high.cholesterol"
  )
covform <- paste0(investigator.specified.covariates, collapse = "+")
out.formula <- as.formula(paste0("outcome", "~", "exposure"))

url <- paste0("https://raw.githubusercontent.com/ehsanx/PlasmodeSim/main/data/hdData/data/scenario/data_1.rds")
data <- readRDS(url(url))
proxy.list <- names(data[, c(grep("^rec", names(data), value = TRUE))])
covarsTfull <- c(investigator.specified.covariates, proxy.list)
Y.form <- as.formula(paste0(c("outcome~ exposure", 
                              covarsTfull), collapse = "+") )
```

```{r initialization before for-loop, warning=FALSE, cache=TRUE}
# 'a' ranges from 1 to 1000 to change the number of datasets loaded in each for-loop
a <- 1000

RD_rf <- data.frame(RD = numeric(), SE = numeric())
```

```{r initialization before for-loop, warning=FALSE, cache=TRUE}
# 'a' ranges from 1 to 1000 to change the number of datasets loaded in each for-loop
a <- 1

RD_rf <- data.frame(RD = numeric(),
                       SE = numeric())
```

```{r scenario: for-loop generating RD & SE results for rf, warning=FALSE, cache=TRUE}
set.seed(42)

# Folder "scenario"
for (i in 1:a) {
  closeAllConnections()
  url <- paste0("https://raw.githubusercontent.com/ehsanx/PlasmodeSim/main/data/hdData/data/scenario/data_", 
                i, 
                ".rds")
  data <- readRDS(url(url))

  # found some id != idx
  data$id <- data$idx

  rf.fit <- randomForest(outcome ~ ., data = data, importance = TRUE)
  # Get important features
  rf.imp <- importance(rf.fit)
  ranked.variables <- rownames(rf.imp)[order(rf.imp[, 1], decreasing = TRUE)]
  proxy_rf <- ranked.variables[ranked.variables %in% proxy.list][1:100]
  proxyform <- paste0(proxy_rf, collapse = "+")
  rhsform <- paste0(c(covform, proxyform), collapse = "+")
  ps.formula <- as.formula(paste0("exposure", "~", rhsform))

  W.out_rf <- weightit(ps.formula, 
                          data = data, 
                          estimand = "ATE",
                          method = "ps")
  #
  data$ps.l <- W.out_rf$ps
  data$w.l <- W.out_rf$weights

  fit.OR_rf <- glm(out.formula,
                      data = data,
                      weights = W.out_rf$weights,
                      family= binomial(link = "logit"))
  fit.RD_rf <- glm(out.formula,
                       data = data,
                       weights = W.out_rf$weights,
                       family = gaussian(link= "identity"))
  sum.RD_rf <- c(summary(fit.RD_rf)$coef["exposure", c("Estimate")], sqrt(sandwich::sandwich(fit.RD_rf)[2,2]))

  RD_rf[i,] <- sum.RD_rf
  rownames(RD_rf)[i] <- paste0("data.", i, "_rf")

  proxy_rf.i <- paste0("proxy_rf.", i)
  assign(proxy_rf.i, proxy_rf)
  sum.RD_rf.i <- paste0("sum.RD_rf.", i)
  assign(sum.RD_rf.i, sum.RD_rf)
}
```

```{r scenarioER: for-loop generating RD & SE results for rf, warning=FALSE, cache=TRUE}
set.seed(42)

# Folder "scenarioER"
for (i in 1:a) {
  closeAllConnections()
  url <- paste0("https://raw.githubusercontent.com/ehsanx/PlasmodeSim/main/data/hdData/data/scenarioER/data_", 
                i, 
                ".rds")
  data <- readRDS(url(url))

  # found some id != idx
  data$id <- data$idx

  rf.fit <- randomForest(outcome ~ ., data = data, importance = TRUE)
  # Get important features
  rf.imp <- importance(rf.fit)
  ranked.variables <- rownames(rf.imp)[order(rf.imp[, 1], decreasing = TRUE)]
  proxy_rf <- ranked.variables[ranked.variables %in% proxy.list][1:100]
  proxyform <- paste0(proxy_rf, collapse = "+")
  rhsform <- paste0(c(covform, proxyform), collapse = "+")
  ps.formula <- as.formula(paste0("exposure", "~", rhsform))

  W.out_rf <- weightit(ps.formula, 
                          data = data, 
                          estimand = "ATE",
                          method = "ps")
  #
  data$ps.l <- W.out_rf$ps
  data$w.l <- W.out_rf$weights

  fit.OR_rf <- glm(out.formula,
                      data = data,
                      weights = W.out_rf$weights,
                      family= binomial(link = "logit"))
  fit.RD_rf <- glm(out.formula,
                       data = data,
                       weights = W.out_rf$weights,
                       family = gaussian(link= "identity"))
  sum.RD_rf <- c(summary(fit.RD_rf)$coef["exposure", c("Estimate")], sqrt(sandwich::sandwich(fit.RD_rf)[2,2]))

  RD_rf[i+a,] <- sum.RD_rf
  rownames(RD_rf)[i+a] <- paste0("ER.data.", i, "_rf")

  proxy_rf.i <- paste0("proxy_rf.ER.", i)
  assign(proxy_rf.i, proxy_rf)
  sum.RD_rf.i <- paste0("sum.RD_rf.ER.", i)
  assign(sum.RD_rf.i, sum.RD_rf)
}
```

```{r scenarioOR: for-loop generating RD & SE results for rf, warning=FALSE, cache=TRUE}
set.seed(42)

# Folder "scenarioOR"
for (i in 1:a) {
  closeAllConnections()
  url <- paste0("https://raw.githubusercontent.com/ehsanx/PlasmodeSim/main/data/hdData/data/scenarioOR/data_", 
                i, 
                ".rds")
  data <- readRDS(url(url))

  # found some id != idx
  data$id <- data$idx

  rf.fit <- randomForest(outcome ~ ., data = data, importance = TRUE)
  # Get important features
  rf.imp <- importance(rf.fit)
  ranked.variables <- rownames(rf.imp)[order(rf.imp[, 1], decreasing = TRUE)]
  proxy_rf <- ranked.variables[ranked.variables %in% proxy.list][1:100]
  proxyform <- paste0(proxy_rf, collapse = "+")
  rhsform <- paste0(c(covform, proxyform), collapse = "+")
  ps.formula <- as.formula(paste0("exposure", "~", rhsform))

  W.out_rf <- weightit(ps.formula, 
                          data = data, 
                          estimand = "ATE",
                          method = "ps")
  #
  data$ps.l <- W.out_rf$ps
  data$w.l <- W.out_rf$weights

  fit.OR_rf <- glm(out.formula,
                      data = data,
                      weights = W.out_rf$weights,
                      family= binomial(link = "logit"))
  fit.RD_rf <- glm(out.formula,
                       data = data,
                       weights = W.out_rf$weights,
                       family = gaussian(link= "identity"))
  sum.RD_rf <- c(summary(fit.RD_rf)$coef["exposure", c("Estimate")], sqrt(sandwich::sandwich(fit.RD_rf)[2,2]))

  RD_rf[i+2*a,] <- sum.RD_rf
  rownames(RD_rf)[i+2*a] <- paste0("OR.data.", i, "_rf")

  proxy_rf.i <- paste0("proxy_rf.OR.", i)
  assign(proxy_rf.i, proxy_rf)
  sum.RD_rf.i <- paste0("sum.RD_rf.OR.", i)
  assign(sum.RD_rf.i, sum.RD_rf)
}
```

```{r read RD & SE results by rf, warning=FALSE, cache=TRUE}
RD_rf
```

```{r check intermediate results for each data by rf, warning=FALSE, cache=TRUE}
# check RD & SE by rf for each data, print:
#   sum.RD_rf.n, sum.RD_rf.ER.n, or sum.RD_rf.OR.n
#   where "n" is a number ranges from 1 to 1000
#   e.g.
sum.RD_rf.1
sum.RD_rf.ER.2

# check proxy selected by rf for each data, print:
#   proxy_rf.n, proxy_rf.ER.n, or proxy_rf.OR.n
#   where "n" is a number ranges from 1 to 1000
#   e.g.
proxy_rf.1
proxy_rf.ER.3
```