### 2024-08-25 Writeup/appendix/Appendix.RMD
writeup

---
title: Supplementary Content
subtitle: Challenges of Univariate Proxy Variable Screening in High-Dimensional Propensity Score Analysis - An Argument for Multivariate Approaches
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    number_sections: yes
  html_document:
    df_print: paged
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
  word_document: default
  bookdown::html_document2:
    citation_package: natbib
    df_print: paged
    fig_caption: yes
    highlight: textmate
    keep_md: yes
    number_sections: yes
    theme: lumen
bibliography: mergedbibliography.bib
biblio-style: plainnat
always_allow_html: yes
header-includes:
- \usepackage{graphicx,longtable,booktabs,here,threeparttable}
- \usepackage{float}
- \usepackage{pdfpages}
- \usepackage{setspace}\singlespacing
- \floatplacement{figure}{H}
- \floatplacement{table}{H}
- \usepackage{xcolor,lipsum}
- \usepackage{caption}
- \captionsetup[table]{labelfont=bf}
- \usepackage[round]{natbib}
- \usepackage{pdflscape}
- \usepackage{lscape}
- \usepackage{etoolbox}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{array}
  \newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
  \newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
  \newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
link-citations: yes
fontsize: 10pt
geometry: margin=.8in
---
\appendix
\counterwithin{figure}{section}
\counterwithin{table}{section}

\edef\tablename{\appendixname~\tablename}
\edef\figurename{\appendixname~\figurename}

\hypertarget{X}{%
\section{X}\label{X}}