### 2024-08-25 Writeup/manuscript.tex

%Version 2.1 April 2023
% See section 11 of the User Manual for version history
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[sn-vancouver,Numbered,lineno,pdflatex]{sn-jnl}

%%%% Standard Packages
%%<additional latex packages if required can be included here>

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%
%%%%

%%%%%=============================================================================%%%%
%%%%  Remarks: This template is provided to aid authors with the preparation
%%%%  of original research articles intended for submission to journals published
%%%%  by Springer Nature. The guidance has been prepared in partnership with
%%%%  production teams to conform to Springer Nature technical requirements.
%%%%  Editorial and presentation requirements differ among journal portfolios and
%%%%  research disciplines. You may find sections in this template are irrelevant
%%%%  to your work and are empowered to omit any such section if allowed by the
%%%%  journal you intend to submit to. The submission guidelines and policies
%%%%  of the journal take precedence. A detailed User Manual is available in the
%%%%  template package for technical guidance.
%%%%%=============================================================================%%%%

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{capt-of}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{threeparttable}
\usepackage{float}
%\floatplacement{figure}{H}
%\floatplacement{table}{H}
\usepackage{lipsum,caption}
\geometry{
  a4paper,
  left=1in,
  right=1in,
  top=1in,
  bottom=1in,
  includeheadfoot
}
\usepackage{setspace}
%\usepackage{lineno}
%\linenumbers
\usepackage{siunitx}
\sisetup{
  mode = match,
  propagate-math-font = true,
  reset-math-version = false,
  reset-text-family = false,
  reset-text-series = false,
  reset-text-shape = false,
  text-family-to-math = true,
  text-series-to-math = true
}
\doublespacing
% Disable explicit page breaks in LaTeX
\let\clearpage\relax

\raggedbottom

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

\title[Deep Learning in Propensity Score Analysis]{Challenges of
Univariate Proxy Variable Screening in High-Dimensional Propensity Score
Analysis - An Argument for Multivariate Approaches}

%%=============================================================%%
%% Prefix	-> \pfx{Dr}
%% GivenName	-> \fnm{Joergen W.}
%% Particle	-> \spfx{van der} -> surname prefix
%% FamilyName	-> \sur{Ploeg}
%% Suffix	-> \sfx{IV}
%% NatureName	-> \tanm{Poet Laureate} -> Title after name
%% Degrees	-> \dgr{MSc, PhD}
%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate}
%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
%%=============================================================%%

\author*[1,2]{\pfx{Dr.} \fnm{Mohammad Ehsanul} \sur{Karim} \dgr{MSc,
PhD}}\email{\href{mailto:ehsan.karim@ubc.ca}{\nolinkurl{ehsan.karim@ubc.ca}}}

\author*[3]{\fnm{Yang} \sur{Lei} }

  \affil*[1]{\orgdiv{School of Population and Public
Health}, \orgname{University of British
Columbia}, \orgaddress{\city{Vancouver}, \country{Canada}, \postcode{V6T
1Z3}, \state{BC}, \street{2206 East Mall}}}
  \affil[2]{\orgname{St.~Paul's
Hospital}, \orgaddress{\city{Vancouver}, \country{Canada}, \postcode{V6Z
1Y6}, \state{BC}, \street{588 - 1081 Burrard Street}}}
  \affil[3]{\orgdiv{Department of Statistics}, \orgname{University of
British
Columbia}, \orgaddress{\city{Vancouver}, \country{Canada}, \postcode{V6T
1Z4}, \state{BC}, \street{Room 3182 Earth Sciences Building, 2207 Main
Mall}}}

\abstract{\textbf{Purpose}:

\textbf{Methods:}

\textbf{Results:}

\textbf{Conclusion:}}

\keywords{Machine learning, Propensity score, Deep learning, Causal
inference}

\pacs[JEL Classification]{C18}
\pacs[MSC Classification]{92D30, 62P10}

\maketitle

\section{Background}\label{background}

\textbf{Aim}:

\section{Methods}\label{methods}

\subsection*{Data and Simulation}\label{data-and-simulation}
\addcontentsline{toc}{subsection}{Data and Simulation}

\textbf{Right Heart Catheterization dataset}:

\textbf{Plasmode simulation}: See Table \ref{table:scenarios} for the
description of the scenarios under consideration.

\begin{table}[ht]
\centering
\caption{Simulation Scenarios for plasmode simulation based on the Right Heart Catheterization (RHC) study.}
\label{table:scenarios}
\begin{tabular}{lcccc}
  \toprule
  \textbf{Plasmode Simulation Scenario} & \textbf{Exposure} & \textbf{Outcome} & \textbf{True} & \textbf{Sample}\\
  \textbf{} & \textbf{Prevalence} & \textbf{Prevalence} & \textbf{Odds Ratio} & \textbf{Size}\\
  \midrule
  (i) Frequent Exposure and Outcome (Base) & 30\% & 30\% & 1 & 3,500 \\
  (ii) Rare Exposure and Frequent Outcome & 5\% & 30\% & 1 & 3,500 \\
  (iii) Frequent Exposure and Rare Outcome & 30\% & 5\% & 1 & 3,500 \\
  \bottomrule
\end{tabular}
\end{table}

\textbf{True data generating mechanism used in plasmode simulation}:

\textbf{Performance measures}: From this simulation, we derived several
performance metrics: (1) bias, (2) average model standard error (SE; the
average of estimated SEs obtained from a model over repeated samples),
(3) empirical SE (the standard deviation of estimated treatment effects
across repeated samples), (4) MSE, (5) coverage probability of 95\%
confidence intervals, (6) bias-eliminated coverage, and (7) Zip plot
\citep{morris2019using, white2023check}.

\subsection*{Estimators under
consideration}\label{estimators-under-consideration}
\addcontentsline{toc}{subsection}{Estimators under consideration}

\section{Results}\label{results}

\section{Real-world analysis}\label{real-world-analysis}

\textbf{Computing time}:

\section{Discussion}\label{discussion}

\textbf{Contextualizing the literature}:

\textbf{Summary of the simulation findings}:

\textbf{Data analysis findings}:

\textbf{Future Direction}:

\textbf{Conclusion}:

\section*{List of abbreviations}\label{list-of-abbreviations}
\addcontentsline{toc}{section}{List of abbreviations}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  MSE - Mean Squared Error
\item
  SE - Standard Error
\item
  PS - Propensity Score
\item
  AE - Autoencoders
\item
  DL - Deep Learning
\item
  MARS - Multivariate Adaptive Regression Splines
\item
  SMD - Standardized Mean Difference
\item
  TMLE - Targeted Maximum Likelihood Estimation
\item
  RHC - Right Heart Catheterization
\item
  SUPPORT - Study to Understand Prognoses and Preferences for Outcomes
  and Risks of Treatments
\end{enumerate}

\section*{Declarations}\label{declarations}
\addcontentsline{toc}{section}{Declarations}

\subsection*{Ethics approval and consent to
participate}\label{ethics-approval-and-consent-to-participate}
\addcontentsline{toc}{subsection}{Ethics approval and consent to
participate}

The analysis conducted on secondary and de-identified data is exempt
from research ethics approval requirements. Ethics for this study was
covered by item 7.10.3 in University of British Columbia's Policy \#89:
Research and Other Studies Involving Human Subjects 19 and Article 2.2
in of the Tri-Council Policy Statement: Ethical Conduct for Research
Involving Humans (TCPS2).

\subsection*{Consent for publication}\label{consent-for-publication}
\addcontentsline{toc}{subsection}{Consent for publication}

\subsection*{Availability of data and
materials}\label{availability-of-data-and-materials}
\addcontentsline{toc}{subsection}{Availability of data and materials}

\subsection*{Competing interests}\label{competing-interests}
\addcontentsline{toc}{subsection}{Competing interests}

Over the past three years, MEK has received consulting fees from Biogen
Inc.~for consulting unrelated to this current work. MEK was previously
supported by the Michael Smith Foundation for Health Research Scholar
award.

\subsection*{Funding}\label{funding}
\addcontentsline{toc}{subsection}{Funding}

This work was supported by MEK's Natural Sciences and Engineering
Research Council of Canada (NSERC) Discovery Grants and Discovery
Accelerator Supplements.

\subsection*{Authors' contributions}\label{authors-contributions}
\addcontentsline{toc}{subsection}{Authors' contributions}

MEK: Conceptualization, Writing -- Original Draft, Review \& Editing YL:
Formal Analysis, Review \& Editing

\subsection*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{subsection}{Acknowledgements}

Not applicable.

\renewcommand\refname{References}
\bibliography{mybibfile.bib}

\end{document}