### 2024-09-03 Writeup/figures/performance_measures_tableER.tex

\begin{table}[!h]
\centering
\begin{tabular}{l|l|l|l|l|l|l}
\hline
  & Bias & Empirical SE & MSE & Model-based SE & Coverage & Bias-eliminated Coverage\\
\hline
\cellcolor{gray!10}{Kitchen sink} & \cellcolor{gray!10}{0.0025 (0.0029)} & \cellcolor{gray!10}{0.0656 (0.0021)} & \cellcolor{gray!10}{0.0043 (0.0005)} & \cellcolor{gray!10}{0.0623 (0.0014)} & \cellcolor{gray!10}{0.9340 (0.0111)} & \cellcolor{gray!10}{0.9320 (0.0113)}\\
\hline
Bross-based hdPS & 0.0035 (0.0027) & 0.0595 (0.0019) & 0.0035 (0.0004) & 0.0586 (0.0013) & 0.9380 (0.0108) & 0.9340 (0.0111)\\
\hline
\cellcolor{gray!10}{Hybrid hdPS} & \cellcolor{gray!10}{0.0049 (0.0025)} & \cellcolor{gray!10}{0.0564 (0.0018)} & \cellcolor{gray!10}{0.0032 (0.0003)} & \cellcolor{gray!10}{0.0564 (0.0012)} & \cellcolor{gray!10}{0.9460 (0.0101)} & \cellcolor{gray!10}{0.9340 (0.0111)}\\
\hline
Genetic Algorithm & 0.0408 (0.0023) & 0.0510 (0.0016) & 0.0043 (0.0003) & 0.0533 (0.0007) & 0.9220 (0.0120) & 0.9420 (0.0105)\\
\hline
\cellcolor{gray!10}{XGBoost} & \cellcolor{gray!10}{0.0259 (0.0023)} & \cellcolor{gray!10}{0.0507 (0.0016)} & \cellcolor{gray!10}{0.0032 (0.0003)} & \cellcolor{gray!10}{0.0531 (0.0009)} & \cellcolor{gray!10}{0.9620 (0.0086)} & \cellcolor{gray!10}{0.9400 (0.0106)}\\
\hline
Random Forest & 0.0127 (0.0027) & 0.0609 (0.0019) & 0.0039 (0.0004) & 0.0593 (0.0014) & 0.9500 (0.0097) & 0.9420 (0.0105)\\
\hline
\cellcolor{gray!10}{Forward Selection} & \cellcolor{gray!10}{0.0108 (0.0024)} & \cellcolor{gray!10}{0.0537 (0.0017)} & \cellcolor{gray!10}{0.0030 (0.0003)} & \cellcolor{gray!10}{0.0555 (0.0011)} & \cellcolor{gray!10}{0.9500 (0.0097)} & \cellcolor{gray!10}{0.9460 (0.0101)}\\
\hline
Backward Elimination & 0.0110 (0.0026) & 0.0576 (0.0018) & 0.0034 (0.0004) & 0.0578 (0.0014) & 0.9520 (0.0096) & 0.9380 (0.0108)\\
\hline
\cellcolor{gray!10}{LASSO} & \cellcolor{gray!10}{0.0060 (0.0026)} & \cellcolor{gray!10}{0.0587 (0.0019)} & \cellcolor{gray!10}{0.0035 (0.0004)} & \cellcolor{gray!10}{0.0580 (0.0014)} & \cellcolor{gray!10}{0.9420 (0.0105)} & \cellcolor{gray!10}{0.9420 (0.0105)}\\
\hline
Elasticnet & 0.0053 (0.0027) & 0.0597 (0.0019) & 0.0036 (0.0004) & 0.0584 (0.0014) & 0.9420 (0.0105) & 0.9380 (0.0108)\\
\hline
\end{tabular}
\end{table}