### 2024-09-13 Writeup/manuscript.tex

%Version 2.1 April 2023
% See section 11 of the User Manual for version history
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[sn-vancouver,Numbered,lineno,pdflatex]{sn-jnl}

%%%% Standard Packages
%%<additional latex packages if required can be included here>

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%
%%%%

%%%%%=============================================================================%%%%
%%%%  Remarks: This template is provided to aid authors with the preparation
%%%%  of original research articles intended for submission to journals published
%%%%  by Springer Nature. The guidance has been prepared in partnership with
%%%%  production teams to conform to Springer Nature technical requirements.
%%%%  Editorial and presentation requirements differ among journal portfolios and
%%%%  research disciplines. You may find sections in this template are irrelevant
%%%%  to your work and are empowered to omit any such section if allowed by the
%%%%  journal you intend to submit to. The submission guidelines and policies
%%%%  of the journal take precedence. A detailed User Manual is available in the
%%%%  template package for technical guidance.
%%%%%=============================================================================%%%%

\usepackage{natbib}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{capt-of}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{threeparttable}
\usepackage{float}
%\floatplacement{figure}{H}
%\floatplacement{table}{H}
\usepackage{lipsum,caption}
\geometry{
  a4paper,
  left=1in,
  right=1in,
  top=1in,
  bottom=1in,
  includeheadfoot
}
\usepackage{setspace}
%\usepackage{lineno}
%\linenumbers
\usepackage{siunitx}
\sisetup{
  mode = match,
  propagate-math-font = true,
  reset-math-version = false,
  reset-text-family = false,
  reset-text-series = false,
  reset-text-shape = false,
  text-family-to-math = true,
  text-series-to-math = true
}
\doublespacing
% Disable explicit page breaks in LaTeX
\let\clearpage\relax

\raggedbottom

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

\title[High-Dimensional Propensity Score]{Is There a Competitive
Advantage to Using Multivariate Statistical or Machine Learning Methods
Over the Bross Formula in the hdPS Framework for Bias and Variance
Estimation?}

%%=============================================================%%
%% Prefix	-> \pfx{Dr}
%% GivenName	-> \fnm{Joergen W.}
%% Particle	-> \spfx{van der} -> surname prefix
%% FamilyName	-> \sur{Ploeg}
%% Suffix	-> \sfx{IV}
%% NatureName	-> \tanm{Poet Laureate} -> Title after name
%% Degrees	-> \dgr{MSc, PhD}
%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate}
%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
%%=============================================================%%

\author*[1,2]{\pfx{Dr.} \fnm{Mohammad Ehsanul} \sur{Karim} \dgr{MSc,
PhD}}\email{\href{mailto:ehsan.karim@ubc.ca}{\nolinkurl{ehsan.karim@ubc.ca}}}

\author[3]{\fnm{Yang} \sur{Lei} }

  \affil*[1]{\orgdiv{School of Population and Public
Health}, \orgname{University of British
Columbia}, \orgaddress{\city{Vancouver}, \country{Canada}, \postcode{V6T
1Z3}, \state{BC}, \street{2206 East Mall}}}
  \affil[2]{\orgname{St.~Paul's
Hospital}, \orgaddress{\city{Vancouver}, \country{Canada}, \postcode{V6Z
1Y6}, \state{BC}, \street{588 - 1081 Burrard Street}}}
  \affil[3]{\orgdiv{Department of Statistics}, \orgname{University of
British
Columbia}, \orgaddress{\city{Vancouver}, \country{Canada}, \postcode{V6T
1Z4}, \state{BC}, \street{Room 3182 Earth Sciences Building, 2207 Main
Mall}}}

\abstract{\textbf{Purpose}: We aims to evaluate various proxy selection
methods within the context of high-dimensional propensity score (hdPS)
analysis. The study focuses on assessing the performance of these
methods, including alternative variable selection approaches, in
selecting proxy variables for confounding adjustment compared to the
traditional hdPS method that is rooted in Bross formula. The goal is to
understand better the performance of these alternative methods in
estimating treatment effects, and identify scenarios in which they may
perform better. \textbf{Methods:} Using data from three cycles of the
National Health and Nutrition Examination Survey (NHANES) spanning
2013-2018, we motivated the study by examining the association between
obesity and diabetes. A plasmode simulation framework based on this data
was employed to mimic real-world data structures. Simulations were
conducted under three scenarios: Frequent Exposure and Outcome, Rare
Exposure and Frequent Outcome, and Frequent Exposure and Rare Outcome.
The performance of a variety of proxy selection methods---including
tree-based methods, LASSO-based methods, and the Genetic Algorithm
(GA)---was evaluated across these scenarios using standard simulation
metrics. \textbf{Results:} XGBoost consistently demonstrated the lowest
MSE and high coverage, making it a reliable method overall, although it
did not always exhibit the lowest bias. In contrast, GA consistently
showed the highest bias and MSE, with lower coverage and greater
variability, indicating its unsuitability for accurate effect
estimation. The kitchen sink model, Bross-based hdPS, and Hybrid hdPS
methods performed moderately well, with low bias and moderate MSE, but
varied in coverage across scenarios. Scenario-specific trends revealed
that rare outcome scenarios yielded lower MSE and better precision,
while rare exposure scenarios were associated with higher bias and MSE.
\textbf{Conclusion:} Aside from GA, the performance of most other
methods in your study appears to be comparable in each scenario, though
with some variations depending on specific metrics. The findings
underscore the importance of selecting appropriate methods for hdPS
analysis based on the characteristics of the data, particularly the
prevalence of exposure and outcome. While XGBoost excelled in overall
accuracy and precision, the choice of method should be tailored to the
specific epidemiological goal to optimize bias, coverage, and precision
in estimating treatment effects.}

\keywords{Machine learning, Propensity score, Deep learning, Causal
inference}

\pacs[JEL Classification]{C18}
\pacs[MSC Classification]{92D30, 62P10}

\maketitle

\section{Background}\label{background}

\textbf{High-dimensional Propensity Score (hdPS) Algorithm}: In
epidemiology, proxy variables are commonly used as substitutes for
confounders that are difficult or impossible to measure directly, such
as socioeconomic status, lifestyle factors, or health behaviors
\citep{vanderweele2019principles}. The high-dimensional propensity score
(hdPS) is a pharmacoepidemiological method designed to reduce
confounding bias in large healthcare databases
\citep{schneeweiss2009high}. Unlike traditional propensity score models
that rely on investigator-specified or manually selected covariates,
hdPS automatically ranks a wide array of proxy variables from healthcare
records---such as diagnosis codes, medications, and procedures---using
the Bross formula \citep{wyss2018erratum, bross1966spurious}. The Bross
formula ranks these variables based on their marginal associations with
both exposure and outcome. These selected proxy variables serve as
surrogates for unmeasured or poorly measured confounders, helping reduce
bias in treatment effect estimates. The hdPS algorithm further refines
the selection by prioritizing variables based on their prevalence and
potential for confounding, defined by their association with both
exposure and outcome \citep{schneeweiss2009high}.

\textbf{Multivariate Machine Learning Extensions}: Although the Bross
formula performs well in certain contexts, it has limitations in
capturing complex interactions, nonlinearities, and higher-order
associations between variables, especially in high-dimensional settings
where it does not account for the multivariate structure of other
covariates \citep{karim2024high, karim2018can}. To address these
limitations, multivariate machine learning methods such as LASSO,
Elastic Net, and Random Forests have been applied within the hdPS
framework. These methods are better suited for high-dimensional data,
where they can more effectively handle complex relationships and improve
the selection of proxy variables, thus enhancing the precision of
treatment effect estimates
\citep{karim2018can, schneeweiss2017variable, franklin2015regularized}.
Simulation studies and empirical research have shown that these machine
learning-based methods, or hybrid approaches combining the Bross formula
with machine learning, can reduce confounding more effectively and
increase efficiency compared to the Bross formula alone in certain
settings
\citep{karim2018can, schneeweiss2017variable, franklin2015regularized}.

\textbf{Assessing the Simulation Performance}: Previous studies have
often focused on Mean Squared Error (MSE) as a primary metric for
evaluating the performance of hdPS and its machine learning extensions
\citep{karim2018can, franklin2015regularized}. However, in
high-dimensional settings with singly robust methods, such as hdPS and
machine learning approaches like LASSO, MSE may not always be the most
reliable measure. MSE combines both bias and variance into a single
metric, which makes interpretation challenging when variance estimation
is unstable---a common issue in these methods. In contrast, coverage,
which measures the proportion of confidence intervals that capture the
true treatment effect, provides a more direct and meaningful assessment
of a model's reliability. In realistic observational studies, where
model misspecification is often inevitable, coverage---along with
related metrics such as bias-eliminated coverage and relative error in
standard error (SE; which compares model-based SE with empirical
SE)---can reveal whether confidence intervals or SEs are too narrow
(underestimating uncertainty) or too wide (overestimating uncertainty).
This insight is crucial in determining whether the model delivers valid
estimates despite misspecification. Even a model with poor MSE but good
coverage may still be valuable, as it produces realistic confidence
intervals. By shifting the focus to coverage, rather than relying solely
on MSE, we can achieve a more comprehensive understanding of method
performance, especially in cases where unstable variance estimation
might distort conclusions drawn from MSE alone.

\textbf{Aim}: This research aims to systematically evaluate and compare
various proxy selection methods within the hdPS framework, using a
diverse range of simulation performance metrics, including bias, MSE,
and coverage. The study focuses on assessing how these alternative
statistical and machine leartning methods perform in selecting proxy
variables for confounding adjustment, compared to the traditional Bross
formula.

\section{Methods}\label{methods}

\subsection*{Data and Simulation}\label{data-and-simulation}
\addcontentsline{toc}{subsection}{Data and Simulation}

\textbf{Motivating Example}: We revisited the association between
obesity and the risk of diabetes using data from three cycles of the
National Health and Nutrition Examination Survey (NHANES) covering the
years 2013-2014, 2015-2016, and 2017-2018 \citep{karim2024high}. To
identify relevant investigator-specified covariates, we constructed a
causal diagram based on literature
\citep{saydah2014trends, liu2013association, kabadi2012joint, ostchega2012abdominal}
and established causal inference principles \citep{greenland1999causal}.
The covariates included in our analysis were carefully selected and
categorized into demographic, behavioral, health history,
access-related, and laboratory variables. While most of these variables
were binary or categorical, the Laboratory variables were continuous.

\textbf{Plasmode simulation}: To rigorously assess the performance of
the methods under consideration, we employed a plasmode simulation
framework, which is particularly well-suited for reflecting real-world
data structures and complexities \citep{franklin2014plasmode}. This
approach was modeled after the analytic dataset derived from NHANES and
involved resampling from the observed covariates and exposure
information (i.e., obesity) without altering them. By mirroring key
aspects of an actual epidemiological study, this simulation framework
offers a substantial advantage over traditional Monte Carlo simulations,
which often rely on hypothetical assumptions.

\textbf{Simulation scenarios under consideration}: Our plasmode
simulation was conducted over 500 iterations. For the base simulation
scenario, we set the prevalence of exposure (obesity) and the event rate
(diabetes) at 30\%, with a true odds ratio (OR) parameter of 1,
corresponding to a risk difference (RD) of 0. Each simulated dataset had
a sample size of 3,000 participants. The description of other scenarios
under consideration is provided in Table \ref{table:scenarios}.

\begin{table}[ht]
\centering
\caption{Overview of Plasmode Simulation Scenarios Reflecting Varying Exposure and Outcome Prevalences Based on National Health and Nutrition Examination Survey (NHANES) Data Cycles (2013-2018)}
\label{table:scenarios}
\begin{tabular}{lcccc}
  \toprule
  \textbf{Plasmode Simulation Scenario} & \textbf{Exposure} & \textbf{Outcome} & \textbf{True} & \textbf{Sample}\\
  \textbf{} & \textbf{Prevalence} & \textbf{Prevalence} & \textbf{Odds Ratio} & \textbf{Size}\\
  \midrule
  (i) Frequent Exposure and Outcome (Base) & 30\% & 30\% & 1 & 3,000 \\
  (ii) Rare Exposure and Frequent Outcome & 5\% & 30\% & 1 & 3,000 \\
  (iii) Frequent Exposure and Rare Outcome & 30\% & 5\% & 1 & 3,000 \\
  \bottomrule
\end{tabular}
\end{table}

\textbf{True Data Generating Mechanism Used in Plasmode Simulation}: The
primary goal of this plasmode simulation study is to evaluate various
variable selection methods under realistic conditions. To achieve this,
we formulated the outcome data based on a specific model specification
that incorporates both exposure and covariates, including
investigator-specified and proxy variables. The model specification
consists of three key components (See Appendices \S A and B for further
details):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \emph{Investigator-Specified Covariates}: We retained the original
  investigator-specified covariates, which were either binary or
  categorical, reflecting how real-world studies typically operate.
\item
  \emph{Transformation of Laboratory Variables}: In real-world studies,
  it is common for analysts to lack precise knowledge of the true model
  specification. To simulate this uncertainty, we transformed the
  continuous laboratory variables using complex functions such as
  logarithmic, exponential, square root, polynomial transformations, and
  interactions. This reflects the challenges analysts face in correctly
  specifying models when dealing with continuous data.
\item
  \emph{Inclusion of Proxy Variables}: Real-world studies often deal
  with unmeasured confounding, which researchers attempt to mitigate by
  adding proxy variables. However, when a large number of proxies are
  added, some may act as noise variables, contributing little or nothing
  to the analysis. To simulate this, we selected only those binary proxy
  covariates (referred to as recurrence covariates in hdPS terminology)
  that had a relative risk (RR) of less than 0.8 or greater than 1.2
  concerning the outcome. Out of 142 proxy covariates, 94 met this
  criterion and were included in calculating a simple comorbidity burden
  measure. The remaining 48 covariates were excluded from this
  calculation and considered noise. This comorbidity burden measure (one
  variable) was then incorporated into our model specification for
  generating the outcome data.
\end{enumerate}

\textbf{Performance Measures}: From this simulation, we derived several
performance metrics to evaluate the effectiveness of the methods under
consideration: (1) bias, (2) average model standard error (SE; the
average of estimated SEs obtained from a model over repeated samples),
(3) empirical SE (the standard deviation of estimated treatment effects
across repeated samples), (4) mean squared error (MSE), (5) coverage
probability of 95\% confidence intervals, (6) bias-corrected coverage,
and (7) Zip plot \citep{morris2019using, white2023check}.

\subsection*{Estimators under
consideration}\label{estimators-under-consideration}
\addcontentsline{toc}{subsection}{Estimators under consideration}

The comparison between the data generation process and the analysis
process reveals two key differences: (i) The data generation used
transformed laboratory variables, whereas the analysis was conducted
using only the original laboratory variables. (ii) The data generation
employed a simple sum of selected proxy variables (sum of 94 proxy
covariates), while the analysis included all proxy variables (142 binary
proxies), with 48 of these acting as noise variables. These differences
help us assess how the proxy variable selection methods handle model
misspecification and the presence of noise variables.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Kitchen sink model}: This is a base model for comparison,
  where no variable selection approaches were used. All
  investigator-selected features and all proxy variables were used to
  model \citep{karim2018can}.
\item
  \textbf{hdPS using Bross formula}: The Bross formula is a statistical
  method used to calculate the bias introduced by not adjusting for a
  covariate \citep{bross1966spurious}. In hdPS analysis, this formula
  was originally applied to each proxy variable to measure and rank the
  potential bias if the covariate were not adjusted for. In our
  analysis, the 100 proxies with the highest bias rankings are selected
  for further modeling \citep{schneeweiss2009high, wyss2018erratum}.
\item
  \textbf{Least Absolute Shrinkage and Selection Operator (LASSO)}:
  LASSO is a variable selection technique that limits the number of
  variables by adding a penalty term to the regression model.
  Cross-validation (CV) is used in LASSO to identify variables with
  non-zero coefficients in the best model by optimizing the penalty
  value
  \citep{franklin2015regularized, schneeweiss2017variable, karim2018can}.
\item
  \textbf{Hybrid of hdPS and LASSO}: Instead of relying solely on LASSO
  for variable selection, a hybrid approach combines the Bross formula
  and LASSO. First, hdPS variables are selected using the hdPS algorithm
  (e.g., the top 100), and then LASSO is applied to further refine the
  selection \citep{karim2018can, franklin2015regularized}.
\item
  \textbf{Elastic Net}: Elastic Net is an extension of LASSO that
  includes an additional penalty term to handle multicollinearity by
  grouping correlated features and selecting the most representative
  ones \citep{karim2018can}.
\item
  \textbf{Random Forest}: The Random Forest (RF) algorithm is an
  ensemble learning method that constructs multiple decision trees to
  perform classification \citep{breiman2001random}. It calculates the
  importance of each proxy variable based on the decrease in impurity or
  Gini importance, providing a ranking of the proxies. The top 100
  variables from this ranking are manually selected for further modeling
  \citep{schneeweiss2017variable}.
\item
  \textbf{XGBoost}: XGBoost is a gradient boosting algorithm used to
  optimize machine learning models \citep{chen2016xgboost}. It builds
  decision trees that make splits based on maximum impurity reduction,
  and it assigns an importance score to each proxy variable by
  calculating the mean decrease in impurity
  \citep{xiao2024interpretable}.
\item
  \textbf{Stepwise}: Stepwise selection is a progressive feature
  selection method that can proceed in two directions---forward or
  backward---based on the maximum adjusted R-squared. We have
  implemented two versions: (a) Forward selection (FS) starts with an
  initial model (e.g., including all investigator-selected features) and
  adds proxies to the model one at a time. (b) Backward elimination (BE)
  starts with a full model (e.g., all investigator-selected features and
  all proxy variables) and removes features one at a time based on their
  contribution to the model.
\item
  \textbf{Genetic algorithm (GA)}: GA is an evolutionary algorithm
  inspired by the theory of natural selection
  \citep{holland1975adaptation}. It operates by evolving offspring from
  a population of the fittest individuals over several generations,
  evaluating and selecting the best combination of features or variables
  that maximize prediction accuracy.
\end{enumerate}

\section{Results}\label{results}

The results for each method under the different scenarios are summarized
below. See Figures \ref{fig:bias-comparison} and
\ref{fig:Coverage-comparison} for an overview of the performance in
terms of bias and coverage, respectively.

\begin{enumerate}
\def\labelenumi{(\roman{enumi})}
\tightlist
\item
  \textbf{Frequent Exposure and Outcome (base) scenario}:
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \emph{Bias}: Bross-based hdPS exhibited the smallest bias (-0.0001),
  and the kitchen sink model (0.0002) was the second. GA shows the
  highest bias (0.0287), indicating a substantial deviation from the
  true effect. Among the other methods, Hybrid hdPS (0.0016), and
  Elastic Net (0.0036) demonstrated low bias. XGBoost (0.0074) had
  slightly higher bias than Random Forest (RF) (0.0034).
\item
  \emph{Coverage}: The coverage for most methods was high, with Hybrid
  hdPS, Forward Selection, Backward Elimination, LASSO, and Elastic Net
  achieving values around 98\%, indicating well-calibrated confidence
  intervals. However, GA had noticeably lower coverage (83.8\%),
  indicating that its confidence intervals might be too narrow or
  biased, potentially missing the true effect. After applying bias
  elimination (as there were substantial bias associated with this
  method), GA's coverage improved to 96\%.
\item
  \emph{Mean Squared Error (MSE)}: XGBoost achieved the lowest MSE
  (0.0006). GA maintained the highest MSE (0.0016), reflecting its
  higher bias and variability. The kitchen sink model (0.0009),
  Bross-based hdPS (0.0008), Hybrid hdPS (0.0008), and Elastic Net
  (0.0009) all had relatively similar and moderate MSE values.
\item
  \emph{Standard Error (SE)}: XGBoost exhibited the lowest Empirical SE
  (0.0229), indicating high precision in its estimates. The kitchen sink
  model had the highest Empirical SE (0.0305), suggesting greater
  variability. Other methods, including GA (0.0274), Hybrid hdPS
  (0.0278), and Bross-based hdPS (0.0287), showed moderate variability.
  LASSO (0.0299) and Elastic Net (0.0294) had slightly higher
  variability. The Model-based SE followed a similar pattern, with
  XGBoost (0.0268) showing the lowest variability and the kitchen sink
  model (0.0333) the highest, indicating less precision in its
  estimates. When comparing relative error in SE estimation, XGBoost
  performed the worst. See Appendix \S C for further details.
\end{enumerate}

\textbf{(ii) Rare Exposure and Frequent Outcome}:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \emph{Bias}: The kitchen sink model showed a relatively low bias
  (0.0025), while GA continued to exhibit the highest bias (0.0408),
  indicating a significant deviation from the true effect. XGBoost had a
  bias of 0.0259, which, while still higher than other methods, but was
  lower than GA. Other methods, such as Bross-based hdPS (0.0035),
  Hybrid hdPS (0.0049), and Elastic Net (0.0053), demonstrated moderate
  bias. Random Forest (RF) (0.0127) and Forward Selection (0.0108) had
  slightly higher bias but remained within an acceptable range.
\item
  \emph{Coverage}: Coverage levels remained high for most methods, with
  XGBoost achieving the highest coverage at 96.2\%, indicating
  well-calibrated confidence intervals despite its higher bias. The GA
  method had lower coverage (92.2\%), suggesting that its confidence
  intervals might be narrower, potentially missing the true effect.
  Other methods such as RF, Forward Selection, Backward Elimination, and
  Hybrid hdPS maintained coverage values around 94-95\%, suggesting
  adequate interval calibration. Bias-eliminated coverage for GA
  improved to 94.2\%, but it was still slightly lower than other methods
  (e.g., forward selection).
\item
  \emph{Mean Squared Error (MSE)}: Forward selection demonstrated the
  lowest MSE (0.0030), and then Hybrid hdPS and XGBoost. The GA method
  had a higher MSE (0.0043), reflecting its substantial bias and
  variability. The kitchen sink model also had an MSE of 0.0043, similar
  to GA, while other methods like Bross-based hdPS (0.0035), RF
  (0.0039), and Elastic Net (0.0036) exhibited moderate MSE values,
  indicating reasonable accuracy.
\item
  \emph{Standard Error (SE)}: The lowest Empirical SE was observed with
  XGBoost (0.0507) and GA (0.0510), reflecting high precision despite
  their higher bias. The kitchen sink model had the highest Empirical SE
  (0.0656), indicating greater variability. Hybrid hdPS (0.0564),
  Bross-based hdPS (0.0595), and RF (0.0609) showed moderate
  variability. Forward Selection (0.0537) and Backward Elimination
  (0.0576) had lower variability compared to the kitchen sink model. In
  terms of Model-based SE, XGBoost (0.0531) and GA (0.0533) continued to
  show low variability, while the kitchen sink model had the highest
  Model-based SE (0.0623), indicating less precision in its estimates.
  When comparing relative error in SE estimation, XGBoost and kitchen
  sink model performed the worst (in other direction).
\end{enumerate}

\textbf{(iii) Frequent Exposure and Rare Outcome}:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \emph{Bias}: In this scenario, the kitchen sink model exhibited a
  moderate negative bias (-0.0093), similar to the Bross-based hdPS
  method (-0.0088). GA showed a significantly higher bias (0.0362),
  indicating a substantial deviation from the true effect. Among other
  methods, XGBoost demonstrated the lowest bias (-0.0061), while methods
  like Hybrid hdPS (-0.0082), Forward Selection (-0.0070), and Backward
  Elimination (-0.0070) had slightly higher but still moderate biases.
  Elastic Net and LASSO both had biases of -0.0079, reflecting slightly
  larger deviations compared to XGBoost but still within acceptable
  limits.
\item
  \emph{Coverage}: Most methods achieved good coverage, with XGBoost,
  RF, and Forward Selection each achieving a coverage rate of 95.4\%,
  indicating well-calibrated confidence intervals. The GA method,
  however, had slightly lower coverage (91.8\%), indicating that its
  confidence intervals might be narrower, potentially excluding the true
  effect. Bross-based hdPS and the kitchen sink model had slightly lower
  coverage values of 93.8\% and 93.4\%, respectively. After accounting
  for bias, the bias-eliminated coverage for most methods, except GA,
  remained high, with values ranging from 98.4\% to 99.0\%, indicating
  that most methods effectively adjusted for bias in their coverage
  estimates. GA's bias-eliminated coverage was lower at 93.4\%,
  reflecting its higher inherent bias.
\item
  \emph{Mean Squared Error (MSE)}: XGBoost exhibited the lowest MSE
  (0.0003). GA had the highest MSE (0.0040), reflecting its substantial
  bias and variability. The kitchen sink model (0.0005), Bross-based
  hdPS (0.0005), and other methods like Hybrid hdPS (0.0004) and Elastic
  Net (0.0005) all had relatively similar MSE values, indicating
  moderate accuracy.
\item
  \emph{Standard Error (SE)}: The lowest Empirical SE was observed with
  XGBoost (0.0152), reflecting high precision in its estimates. The GA
  method exhibited the highest Empirical SE (0.0523), indicating greater
  variability and less precision. Methods such as Hybrid hdPS (0.0184),
  Forward Selection (0.0187), and Elastic Net (0.0203) showed moderate
  variability, while Bross-based hdPS (0.0206) and the kitchen sink
  model (0.0212) had slightly higher variability. In terms of
  Model-based SE, XGBoost (0.0179) again showed the lowest variability,
  consistent with its low Empirical SE, indicating that it provided the
  most stable estimates. The kitchen sink model had a slightly higher
  Model-based SE (0.0219), indicating less precision in its estimates.
  When comparing relative error in SE estimation, XGBoost performed the
  worst.
\end{enumerate}

\begin{figure}[th]

{\centering \includegraphics[width=1\linewidth,]{figures/metric_comparison_Bias}

}

\caption{Comparison of Bias Across Different Methods in hdPS Analysis\label{fig:bias-comparison}}\label{fig:unnamed-chunk-1}
\end{figure}

\begin{figure}[th]

{\centering \includegraphics[width=1\linewidth,]{figures/metric_comparison_Coverage}

}

\caption{Comparison of Coverage Probability Across Different Methods in hdPS Analysis\label{fig:Coverage-comparison}}\label{fig:unnamed-chunk-2}
\end{figure}

\section{Real-world analysis}\label{real-world-analysis}

\textbf{Summary results}: The dataset comprises 7,585 individuals. Among
these, the prevalence of the exposure is 48.8\%, while the prevalence of
the outcome is 23.7\%.

See Figure \ref{fig:comparison-plot-updated} for the results from
analyzing the NHANES (2013-2018) dataset. The methods are arranged
according to the number of selected proxy variables. Among all variable
selection algorithms, Random Forest (RF) and XGBoost demonstrate the
highest odds ratios (ORs), with values of 1.59 and 1.56, respectively.
The ORs for the remaining methods cluster around 1.52. Additionally,
with the exception of RF and the Bross formula hdPS, a general pattern
emerges where methods selecting a larger number of proxy variables yield
lower ORs. For risk difference (RD), RF and XGBoost also exhibit the
highest value 0.08, while the remaining methods converge around 0.077.
The trend observed in the OR results appears to persist in the RD
analysis.

\begin{figure}[th]

{\centering \includegraphics[width=1\linewidth,]{manuscript_files/figure-latex/unnamed-chunk-3-1}

}

\caption{Figure presenting a comparison of Risk Differences (RD) and Odds Ratios (OR) with 95\% confidence intervals for different methods used to evaluate the association between obesity and diabetes risk. The analysis is based on data from the National Health and Nutrition Examination Survey (NHANES) for the years 2013-2018. Methods are arranged by the number of variables used in the models.\label{fig:comparison-plot-updated}}\label{fig:unnamed-chunk-3}
\end{figure}

Table \ref{tab:method-comparison-updated} presents a pairwise comparison
of the number of proxy features shared between different variable
selection methods used in the analysis. Each cell in the table indicates
the count of common proxy variables selected by the method in the
corresponding row and column. The diagonal cells, where the row and
column methods are the same, represent the total number of proxy
variables selected exclusively by each method. The first column displays
the number of proxy variables shared between each method and the kitchen
sink (KS) model. Given that the KS model includes all proxy variables,
the values in the first column are identical to those in the diagonal
cells for each row, which also presents the total number of proxy
variables selected by the method in the corresponding row.

\begin{table}[htbp]
\centering
\caption{Comparison of variable overlap of selected proxies across different methods used to evaluate the association between obesity and diabetes}
\label{tab:method-comparison-updated}
\begin{tabular}{lcccccccccc}
\toprule
 & \textbf{KS} & \textbf{Bross} & \textbf{Hybrid} & \textbf{LASSO} & \textbf{EN} & \textbf{RF} & \textbf{XGB} & \textbf{FS} & \textbf{BE} & \textbf{GA} \\
\midrule
\textbf{Kitchen sink (KS)} & 142 & & & & & & & & & \\
\textbf{Bross formula} & 100 & 100 & & & & & & & & \\
\textbf{Hybrid (Bross and LASSO)} & 49 & 49 & 49 & & & & & & & \\
\textbf{LASSO} & 60 & 47 & 47 & 60 & & & & & & \\
\textbf{Elastic Net (EN)} & 69 & 54 & 48 & 60 & 69 & & & & & \\
\textbf{Random Forest (RF)} & 100 & 71 & 38 & 46 & 52 & 100 & & & & \\  
\textbf{XGBoost (XGB)} & 48 & 38 & 24 & 28 & 30 & 37 & 48 & & & \\
\textbf{Forward selection (FS)} & 59 & 45 & 41 & 51 & 54 & 45 & 25 & 59 & & \\
\textbf{Backward elimination (BE)} & 59 & 45 & 41 & 51 & 54 & 45 & 25 & 59 & 59 & \\
\textbf{Genetic algorithm (GA)} & 64 & 44 & 28 & 36 & 40 & 49 & 25 & 35 & 35 & 64 \\
\bottomrule
\end{tabular}
\end{table}

Table \ref{tab:proxy-common-pct-bross} presents a comparison of the
count and percentage of common proxy variables selected by different
methods in comparison with the Bross formula hdPS. The first column
shows the total number of proxy variables selected by each method, while
the second column lists the number of common features selected by both
the respective method and the Bross formula. The third column reports
the percentage of common features out of the total number of features
selected by each method.

We observe that, aside from the Bross formula hdPS and Random Forest
(RF) models, where the number of proxies was manually set to 100, and
the kitchen sink (KS) model, which includes all proxies by design, the
number of proxy variables selected by other models ranges between 49 and
69. As expected, the hybrid method combining Bross and LASSO hdPS
selects exactly 49 proxy variables, all of which are selected by both
methods, resulting in a common feature rate of 1.00. For other models,
the common feature percentage is generally clustered around \(74\%\),
with XGBoost showing the highest common percentage at \(79\%\), while
the Genetic Algorithm (GA) displays the lowest common percentage at
\(69\%\).

\begin{table}[htbp]
\centering
\caption{Comparison of the count and percentage of proxy variables selected by each methods in common with that by the Bross formula hdPS}
\label{tab:proxy-common-pct-bross}
\begin{tabular}{lccc}
\toprule
\textbf{Method} & \textbf{Total Count} & \textbf{Common Count} & \textbf{Rate in Common} \\ 
\midrule
\textbf{Kitchen sink} & 142 & - & - \\
\textbf{Bross formula} & 100 & - & - \\
\textbf{Hybrid (Bross and LASSO)} & 49 & 49 & 1.00 \\
\textbf{LASSO} & 60 & 47 & 0.78 \\
\textbf{Elastic Net} & 69 & 54 & 0.78 \\
\textbf{Random Forest (RF)} & 100 & 71 & 0.71 \\  
\textbf{XGBoost} & 48 & 38 & 0.79 \\
\textbf{Forward selection (FS)} & 59 & 45 & 0.76 \\
\textbf{Backward elimination (BE)} & 59 & 45 & 0.76 \\
\textbf{Genetic algorithm (GA)} & 64 & 44 & 0.69 \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Computing time}:

Figure \ref{fig:computing-time} presents the computing time for each
method. All methods, aside from RF and GA, exhibit relatively fast
computing times. RF and GA are generally much slower.

\begin{figure}[th]

{\centering \includegraphics[width=1\linewidth,]{manuscript_files/figure-latex/unnamed-chunk-4-1}

}

\caption{Computing time for the real-world analysis for each algorithm under consideration. The analysis is based on data from the National Health and Nutrition Examination Survey (NHANES) for the years 2013-2018.\label{fig:computing-time}}\label{fig:unnamed-chunk-4}
\end{figure}

\section{Discussion}\label{discussion}

\subsection{Summary of the simulation
findings}\label{summary-of-the-simulation-findings}

\textbf{Comparison of methods}:

Across the three scenarios---Frequent Exposure and Outcome, Rare
Exposure and Frequent Outcome, and Frequent Exposure and Rare
Outcome---XGBoost consistently exhibited the lowest Mean Squared Error
(MSE) and high coverage, making it one of the most reliable methods
overall. However, XGBoost did not always have the lowest bias; in fact,
methods such as the kitchen sink model, Bross-based hdPS, and Hybrid
hdPS often showed lower bias. The GA, in contrast, consistently
exhibited the highest bias and MSE, along with lower coverage and
greater variability, indicating it was less reliable for accurate effect
estimation. The kitchen sink model, while demonstrating low bias and
reasonable MSE, generally had higher variability and lower precision
compared to other methods like XGBoost. Methods such as Bross-based
hdPS, Hybrid hdPS, and Elastic Net performed moderately well across all
scenarios, balancing bias, coverage, and MSE, but they did not
outperform XGBoost in overall accuracy and precision.

\textbf{Comparison of scenarios}:

Scenarios with rare exposure tended to produce higher bias, particularly
for methods like GA and XGBoost, while frequent outcomes generally led
to lower bias. Overcoverage was common in the frequent exposure and
outcome scenario, while the other scenarios tended to have more balanced
or slightly under-coverage. Frequent exposure scenarios had higher
relative error, indicating more difficulty in precisely estimating
effects compared to rare exposure scenarios, which had lower relative
error. Rare outcome scenarios tended to have the lowest MSE, indicating
more precise effect estimates under these conditions, whereas rare
exposure scenarios had higher MSE, reflecting the challenges of
estimating effects when exposure is uncommon.

\subsection{Contextualizing the
literature}\label{contextualizing-the-literature}

\textcolor{red}{*need to write what hdps offers in the literature; what additional selection criteria was used so far*}

\subsection{Data analysis findings}\label{data-analysis-findings}

\textcolor{red}{*summarize real data analysis finding in 1 para*}

\subsection{Future Direction}\label{future-direction}

\textcolor{red}{*TMLE*}

\subsection{Conclusion}\label{conclusion}

In conclusion, this analysis highlights the importance of carefully
selecting appropriate methods for hdPS analysis based on the specific
characteristics of the data, particularly the prevalence of exposure and
outcome. While XGBoost consistently demonstrated strong performance in
terms of MSE and coverage, it did not always have the lowest bias,
suggesting that it may be most suitable when precision is prioritized
over bias minimization. GA, despite its potential, showed significant
limitations with consistently high bias and MSE, making it less reliable
for accurate effect estimation. The kitchen sink model, Bross-based
hdPS, and Hybrid hdPS methods provided a balanced approach, often
delivering low bias and moderate MSE, but with variability in coverage
depending on the scenario. Scenario-specific trends revealed that rare
outcomes generally yielded lower MSE and better precision, while rare
exposures were associated with higher bias and MSE, emphasizing the
challenges of accurately estimating effects in such contexts.
Ultimately, the findings underscore the need to tailor method selection
to the epidemiological scenario at hand, ensuring that the chosen
approach aligns with the specific goals and challenges of the analysis.

\section*{List of abbreviations}\label{list-of-abbreviations}
\addcontentsline{toc}{section}{List of abbreviations}

\begin{itemize}
\tightlist
\item
  hdPS: High-dimensional Propensity Score
\item
  NHANES: National Health and Nutrition Examination Survey
\item
  OR: Odds Ratio
\item
  RD: Risk Difference
\item
  SE: Standard Error
\item
  MSE: Mean Squared Error
\item
  \textcolor{blue}{KS: Kitchen Sink}
\item
  LASSO: Least Absolute Shrinkage and Selection Operator
\item
  \textcolor{blue}{ElasticNet / EN}: A regularized regression method
  that combines LASSO and Ridge regression
\item
  RF: Random Forest
\item
  XGBoost: Extreme Gradient Boosting
\item
  FS: Forward Selection
\item
  BE: Backward Elimination
\item
  GA: Genetic Algorithm
\item
  CV: Cross-Validation
\item
  RR: Relative Risk
\end{itemize}

\section*{Declarations}\label{declarations}
\addcontentsline{toc}{section}{Declarations}

\subsection*{Ethics approval and consent to
participate}\label{ethics-approval-and-consent-to-participate}
\addcontentsline{toc}{subsection}{Ethics approval and consent to
participate}

The analysis conducted on secondary and de-identified data is exempt
from research ethics approval requirements. Ethics for this study was
covered by item 7.10.3 in University of British Columbia's Policy \#89:
Research and Other Studies Involving Human Subjects 19 and Article 2.2
in of the Tri-Council Policy Statement: Ethical Conduct for Research
Involving Humans (TCPS2).

\subsection*{Consent for publication}\label{consent-for-publication}
\addcontentsline{toc}{subsection}{Consent for publication}

\subsection*{Availability of data and
materials}\label{availability-of-data-and-materials}
\addcontentsline{toc}{subsection}{Availability of data and materials}

\subsection*{Competing interests}\label{competing-interests}
\addcontentsline{toc}{subsection}{Competing interests}

Over the past three years, MEK has received consulting fees from Biogen
Inc.~for consulting unrelated to this current work. MEK was previously
supported by the Michael Smith Foundation for Health Research Scholar
award.

\subsection*{Funding}\label{funding}
\addcontentsline{toc}{subsection}{Funding}

This work was supported by MEK's Natural Sciences and Engineering
Research Council of Canada (NSERC) Discovery Grants and Discovery
Accelerator Supplements.

\subsection*{Authors' contributions}\label{authors-contributions}
\addcontentsline{toc}{subsection}{Authors' contributions}

MEK: Conceptualization, Writing -- Original Draft, Review \& Editing YL:
Formal Analysis, Review \& Editing

\subsection*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{subsection}{Acknowledgements}

Not applicable.

\renewcommand\refname{References}
\bibliography{mergedbibliography.bib}

\end{document}