### 2024-09-14 Writeup/appendix/Appendix.RMD

---
title: Supplementary Content
subtitle: Is There a Competitive Advantage to Using Multivariate Statistical or Machine Learning Methods Over the Bross Formula in the hdPS Framework for Bias and Variance Estimation?
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    number_sections: yes
  html_document:
    df_print: paged
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
  word_document: default
  bookdown::html_document2:
    citation_package: natbib
    df_print: paged
    fig_caption: yes
    highlight: textmate
    keep_md: yes
    number_sections: yes
    theme: lumen
bibliography: mergedbibliography.bib
biblio-style: plainnat
always_allow_html: yes
header-includes:
- \usepackage{graphicx,longtable,booktabs,here,threeparttable}
- \usepackage{float}
- \usepackage{pdfpages}
- \usepackage{setspace}\singlespacing
- \floatplacement{figure}{H}
- \floatplacement{table}{H}
- \usepackage{xcolor,lipsum}
- \usepackage{caption}
- \captionsetup[table]{labelfont=bf}
- \usepackage[round]{natbib}
- \usepackage{pdflscape}
- \usepackage{lscape}
- \usepackage{etoolbox}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{array}
  \newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
  \newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
  \newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
link-citations: yes
fontsize: 10pt
geometry: margin=.8in
---
\appendix
\counterwithin{figure}{section}
\counterwithin{table}{section}

\edef\tablename{\appendixname~\tablename}
\edef\figurename{\appendixname~\figurename}

# Variables Used for Plasmode Simulation Data Generation

1. Original demographic variables (8)

- age, 
- sex, 
- education, 
- race, 
- marital status, 
- income, 
- country where born, 
- survey cycle

2. Original behaviour variables (5)

- smoking, 
- diet, 
- high cholesterol, 
- physical activity, 
- sleep

3. Original health history / access variables (2)

- diabetes family history, 
- medical access

4. Transformed lab variables (6) (complex forms) based on original lab variables: uric acid, protein, bilirubin, phosphorus, sodium, potassium, globulin, calcium, systolic blood pressure, diastolic blood pressure.

- Tranfored.var.1 = log(globulin)
- Tranfored.var.2 = protein*calcium
- Tranfored.var.3 = diastolicBP/systolicBP)^2
- Tranfored.var.4 = sqrt(uric acid+bilirubin)/2
- Tranfored.var.5 = phosphorus^2/(sodium*potassium)
- Tranfored.var.6 = log(systolicBP+10)

5. Count based prescription codes (1) (proxies of comorbidity)

Simple count (1 variable) = sum of selected ICD-10 CM codes (converted to recurrence covariates) who had less than 0.8 or greater than 1.2 compared to the outcome = $\sum_{s=1}^{94} R_s$

# True Outcome Model for Plasmode Simulation Data Generation

Diabetes (outcome) =  Obese (exposure) + demographic/behaviour/health history variables + transformed lab variables + simple count with selected ICD-10 codes

# Standard Error comparison

```{r, echo=FALSE, out.width="100%", fig.cap="Standard Error Comparison for Different Methods (Overall) when outcome and exposure are frequent.", fig.align="center", fig.label="fig:se-comparison-overall"}
knitr::include_graphics("se_comparison_plot.png")
```

```{r, echo=FALSE, out.width="100%", fig.cap="Standard Error Comparison for Different Methods (Overall) when outcome is frequent but exposure is rare.", fig.align="center", fig.label="fig:se-comparison-overall"}
knitr::include_graphics("se_comparison_plotER.png")
```

```{r, echo=FALSE, out.width="100%", fig.cap="Standard Error Comparison for Different Methods (Overall) when outcome is rare but exposure is frequent", fig.align="center", fig.label="fig:se-comparison-overall"}
knitr::include_graphics("se_comparison_plotOR.png")
```